### 2024-03-29 paper/paper.qmd
Basic setup

---
title: "Title"
subtitle: "Subtitle"
author: Kaavya Kalani
thanks: "Code and data supporting this analysis is available at: "
date: today
date-format: long
abstract: "Abstract"
format: pdf
number-sections: true
toc: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

```

# Introduction

The paper is further organized into four sections: Data, Model, Results, and Discussion. In the Data section, we discuss how the datasets to be used for the analysis were obtained and pre-processed. We will explain the variables of interest in the datasets for the analysis. The Model section describes the model being used for the analysis. The Results section will then highlight and discuss the trends and associations found during the analysis. Lastly, the Discussion section will talk about some interesting trends found in Results in depth, link it to the real world and also highlight the weaknesses and future of our analysis.

# Data

# Model

# Results

# Discussion

# Appendix

# References

### 2024-03-29 scripts/00-simulate_data.R

@@ -1,0 +1,9 @@
+#### Preamble ####
+# Purpose: Simulated the survey and poststratification datasets
+# Author: Kaavya Kalani
+# Date: XX March 2024
+# Contact: kaavya.kalani@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: None
+
+#### Workspace setup ####

### 2024-03-29 scripts/01-download_data.R

@@ -1,0 +1,9 @@
+#### Preamble ####
+# Purpose: Downloads and saves the data from XX
+# Author: Kaavya Kalani
+# Date: XX March 2024
+# Contact: kaavya.kalani@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: 
+
+#### Workspace setup ####

### 2024-03-29 scripts/02-data_cleaning.R

@@ -1,0 +1,9 @@
+#### Preamble ####
+# Purpose: Cleans the raw survey and raw poststratification data
+# Author: Kaavya Kalani
+# Date: XX March 2024
+# Contact: kaavya.kalani@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: 
+
+#### Workspace setup ####

### 2024-03-29 scripts/03-test_data.R

@@ -1,0 +1,9 @@
+#### Preamble ####
+# Purpose: Tests survey and postratification data
+# Author: Kaavya Kalani
+# Date: 9 March 2024
+# Contact: kaavya.kalani@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: Run 01-data_cleaning.R
+
+#### Workspace setup ####

### 2024-03-29 scripts/04-model.R

@@ -1,0 +1,9 @@
+#### Preamble ####
+# Purpose: XX model
+# Author: Kaavya Kalani
+# Date: XX March 2024
+# Contact: kaavya.kalani@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: Run 01-data_cleaning.R
+
+#### Workspace setup ####

### 2024-03-29 scripts/05-poststratification.R

@@ -1,0 +1,9 @@
+#### Preamble ####
+# Purpose: Poststratification for the model
+# Author: Kaavya Kalani
+# Date: XX March 2024
+# Contact: kaavya.kalani@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: Run 04-model.R
+
+#### Workspace setup ####

### 2024-04-13 paper/paper.qmd
Paper structure layed out and R scripts created

@@
 ---
-title: "Title"
-subtitle: "Subtitle"
+title: "Himalayan Expeditions Success Analysis "
+subtitle: "Analysis using data from XXX leveraging Bayesian Logistic Regression"
 author: Kaavya Kalani
-thanks: "Code and data supporting this analysis is available at: "
+thanks: "Code and data supporting this analysis is available at: https://github.com/kaavyakalani26/himalayan-expeditions-analysis"
 date: today
@@
 
+library(broom.mixed)
+library(tidyverse)
+library(dplyr)
+library(knitr)
+library(ggplot2)
+library(modelsummary)
+library(here)
+library(kableExtra)
+
+expeditions_model <- readRDS(here("models/single_bay.rds"))
+analysis_data <- read.csv(here("data/analysis_data/expeditions.csv"))
 ```
@@
 
-The paper is further organized into four sections: Data, Model, Results, and Discussion.
-In the Data section, we discuss how the datasets to be used for the analysis were obtained and pre-processed.
-We will explain the variables of interest in the datasets for the analysis.
-The Model section describes the model being used for the analysis.
-The Results section will then highlight and discuss the trends and associations found during the analysis.
-Lastly, the Discussion section will talk about some interesting trends found in Results in depth, link it to the real world and also highlight the weaknesses and future of our analysis.
+1) broader context to motivate; 
+2) some detail about what the paper is about; 
+3) a clear gap that needs to be filled; 
+4) what was done; 
+5) what was found; 
+6) why it is important;
+7) Estimand
 
-# Data
+(Likely 3 or 4 paragraphs, or 10 per cent of total.)
 
-# Model
+The paper is further organized into four sections.
+@sec-data discusses how the dataset to be used for the analysis was obtained and pre-processed.
+I will explain the variables of interest in the dataset for the analysis.
+@sec-model describes the model being used for the analysis.
+@sec-results then highlights and discusses the trends and associations found during the analysis.
+Lastly,@sec-discussion talks about some interesting trends found in @sec-results in depth, link it to the real world and also highlight the weaknesses and future of my analysis.
 
-# Results
+# Data {#sec-data}
+For this analysis, we have used combined three datasets into one, which is used for analysis.
+The datasets were cleaned and analysed using the statistical programming software `R` [@citeR] along with the help `tidyverse` [@citeTidyverse], `knitr` [@citeKnitr], `ggplot2` [@citeGgplot], `here` [@citeHere], `dplyr` [@citeDplyr], `rstanarm` [@citeRstanarm], `broom.mixed` [@citeBroomMixed], `modelsummary` [@citeModelSummary] and `kableExtra` [@citeKableExtra].
 
-# Discussion
+## Analysis Dataset
+The analysis data is from XXXX. This is how it was collected.
+
+A person becomes an entry in this dataset if XXX. (Measurement)
+
+The data from datasets for XXX are combined to form our anaysis dataset. Among the overall range of variables available, we chose XXX to be included in our analysis dataset.
+
+- Describe the variables
+
+Out of these, we will be using XXX in our model as the independent variables and XXX as the dependent variable.
+
+# Model {#sec-model}
+I used a Bayesian Logistic Regression model to do XXX. Logistic regression is a method used for binary classification to predict the probability of a categorical dependent variable.
+
+For my analysis, a logistic regression model will be first used to model XXX. The model will be based on XXX independent demographic variables: XXX. (Use backticks)
+
+The logistic regression model I will be using is:
+\begin{equation}
+\log\left(\frac{\hat{p}}{1 - \hat{p}}\right) = \beta_0 + \beta_1 \times \text{height} + \beta_2 \times \text{sex} + \beta_3 \times \text{age} + \beta_4 \times \text{seasons} + \beta_5 \times \text{solo}
+\end{equation}
+
+$$
+\begin{aligned}
+\beta_0 & \sim \mbox{Normal}(0, 2.5)\\
+\beta_1 & \sim \mbox{Normal}(0, 2.5)\\
+\beta_2 & \sim \mbox{Normal}(0, 2.5)\\
+\beta_3 & \sim \mbox{Normal}(0, 2.5)\\
+\beta_4 & \sim \mbox{Normal}(0, 2.5)\\
+\beta_5 & \sim \mbox{Normal}(0, 2.5)
+\end{aligned}
+$$
+where,
+
+- $\hat{p}$ represents the probability that someone will successfully complete the peak they are on the expedition for.
+- $\beta_0$ represents the intercept term of this logistical regression.
+It is the probability that someone will successfully complete the peak they are on the expedition for if the predictors' values are zero
+- $\beta_1$ is the coefficient corresponding to height of the peak
+- $\beta_2$ is the coefficient corresponding to sex
+- $\beta_3$ is the coefficients corresponding to age
+- $\beta_4$ is the coefficients corresponding to seasons
+- $\beta_5$ is the coefficients corresponding to solo
+
+In my model, normal priors with a mean of 0 and a standard deviation of 2.5 are used for both the coefficients and the intercept.
+Setting the mean of the priors to 0 implies that there is no expectation of a particular direction or magnitude for the coefficients or intercept.
+I chose this as I have no expectation of the same.
+The standard deviation of 2.5 reflects the uncertainty or variability in the prior beliefs.
+I chose a moderately wide prior to allow for a reasonable amount of uncertainty.
+
+The chosen priors allow the data to largely determine the posterior distribution as they are relatively non-informative.
+They don't heavily influence the results unless the data provide strong evidence to the contrary.
+
+The use of moderately wide priors can also help regularize the model, preventing overfitting and providing more stable estimates, particularly when dealing with limited data.
+
+```{r}
+#| echo: false
+#| warning: false
+#| message: false
+#| label: tbl-model-coefficients
+#| tbl-cap: Summary of the model
+
+# Create model summary
+# model_summary <- modelsummary(expeditions_model)
+
+# Extract coefficients
+coefficients <- broom::tidy(expeditions_model, conf.int = TRUE)
+coefficients |> kable(digits = 2)
+```
+@tbl-model-coefficients shows the coefficients for my Bayesian model along with the standard error and the 95% credible interval.
+The standard error (SE) is a measure of the precision with which a sample statistic estimates a population parameter.
+It quantifies the variability of sample statistics around the population parameter.
+A 95% credible interval means that there is a 95% probability that the true parameter lies within the interval, given the observed data and the model assumptions.
+
+```{r}
+#| echo: false
+#| warning: false
+#| message: false
+#| label: fig-model-coefficients
+#| fig-cap: Coefficients of the model
+
+ggplot(coefficients, aes(x = estimate, y = term)) +
+  geom_point() + # Coefficient points
+  geom_errorbarh(aes(xmin = conf.low, xmax = conf.high)) + # Confidence/Credible interval bars
+  geom_vline(xintercept = 0, linetype = "dashed", color = "red") + # Add zero line
+  labs(x = "Coefficient Estimate", y = "Predictor Variable") + # Axis labels
+  theme_minimal() # Minimal theme for better readability
+``` 
+@fig-model-coefficients illustrates the coefficients and their associated 95% credible intervals for the predictor variables in the Bayesian model.
+Each point represents the estimated coefficient for a predictor variable, while the horizontal lines depict the credible interval around the estimate.
+Variables with coefficients to the right of zero indicate a positive association with the outcome variable, suggesting that an increase in the predictor variable corresponds to an increase in the outcome variable.
+Conversely, coefficients to the left of zero indicate a negative association, implying that an increase in the predictor variable is associated with a decrease in the outcome variable.
+These insights can help in understanding the direction and magnitude of the relationships between predictor variables and the outcome in the Bayesian model.
+
+# Results {#sec-results}
+
+```{r}
+#| echo: false
+#| warning: false
+#| message: false
+#| label: fig-results
+#| fig-cap: Results1
+
+# Create a data frame with all combinations of sex and seasons
+sex_levels <- c('M', 'F')  # Replace with actual categories
+seasons_levels <- unique(analysis_data$seasons)  # Extract unique seasons from your analysis data
+
+new_data <- expand.grid(
+  sex = sex_levels,
+  seasons = seasons_levels
+)
+
+# Predicting success based on sex and seasons
+predictions <- predict(expeditions_model, newdata = new_data, type = "response")
+
+# Create a data frame to display the predictions
+prediction_df <- data.frame(
+  sex = new_data$sex, 
+  seasons = new_data$seasons, 
+  success_probability = predictions
+)
+
+ggplot(prediction_df, aes(x = sex, y = success_probability, fill = sex)) +
+  geom_bar(stat = "identity", position = "dodge") +
+  labs(title = "Predicted Success Probability by Gender",
+       x = "Sex",
+       y = "Success Probability") +
+  theme_minimal()
+
+```
+
+# Discussion {#sec-discussion}
+1) What is done in this paper? 
+2) What is something that we learn about the world?
+3) What are some weaknesses of what was done? What is left to learn or how should we proceed in the future?
+
+## Weaknesses and Limitations
+
+## Future Directions
 
@@
 
+## Analysis dataset
+Here is a glimpse of the dataset used for analysis
+
+```{r}
+#| echo: false
+#| warning: false
+#| message: false
+#| label: tbl-survey
+#| tbl-cap: Analysis dataset
+#| 
+head(analysis_data)  %>%
+  kable(digits = 2)
+```
+
 # References

### 2024-04-13 scripts/00-install-packages.R

@@ -1,0 +1,26 @@
+#### Preamble ####
+# Purpose: Install all necessary packages for this paper's replication
+# Author: Kaavya Kalani
+# Date:6 April 2024
+# Contact: kaavya.kalani@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: None
+
+# List of packages to install
+packages <- c("tidyverse", 
+              "rstanarm", 
+              "broom.mixed", 
+              "dplyr", 
+              "knitr", 
+              "ggplot2", 
+              "modelsummary", 
+              "here", 
+              "kableExtra")
+
+# Install packages if they are not already installed
+for (package in packages) {
+  if (!requireNamespace(package, quietly = TRUE)) {
+    install.packages(package, dependencies = TRUE)
+  }
+}
+

### 2024-04-13 scripts/01-data_cleaning.R

@@ -1,9 +1,0 @@
-#### Preamble ####
-# Purpose: Cleans the raw survey and raw poststratification data
-# Author: Kaavya Kalani
-# Date: XX March 2024
-# Contact: kaavya.kalani@mail.utoronto.ca
-# License: MIT
-# Pre-requisites: 
-
-#### Workspace setup ####

### 2024-04-13 scripts/02-data_cleaning.R

@@ -1,0 +1,32 @@
+#### Preamble ####
+# Purpose: Cleans the raw data
+# Author: Kaavya Kalani
+# Date:6 April 2024
+# Contact: kaavya.kalani@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: None
+
+#### Workspace setup ####
+library(tidyverse)
+
+#### Load and clean data ####
+peaks <- read.csv("data/raw_data/peaks.csv")
+expeditions <- read.csv("data/raw_data/expeditions.csv")
+members <- read.csv("data/raw_data/members.csv")
+
+merged_data1 <- merge(expeditions, members, by = "expedition_id")
+merged_data1 <- merged_data1 %>%
+  select(peak_id.x, season.x, sex, age, success, solo, died, members) %>%
+  rename(peak_id = peak_id.x)
+
+peaks <- peaks %>%
+  select(peak_id, height_metres)
+
+merged_data2 <- merge(merged_data1, peaks, by = "peak_id")
+main_data <- merged_data2 %>%
+  filter(!is.na(sex), !is.na(age), members > 0) %>%
+  select(peak_id, height_metres, season.x, sex, age, success, solo, died, members) %>%
+  rename(height = height_metres, seasons = season.x)
+
+#### Save the cleaned data ####
+write.csv(main_data, "data/analysis_data/expeditions.csv", row.names = FALSE)

### 2024-04-13 scripts/03-model.R

@@ -1,9 +1,0 @@
-#### Preamble ####
-# Purpose: XX model
-# Author: Kaavya Kalani
-# Date: XX March 2024
-# Contact: kaavya.kalani@mail.utoronto.ca
-# License: MIT
-# Pre-requisites: Run 01-data_cleaning.R
-
-#### Workspace setup ####

### 2024-04-13 scripts/03-test_data.R

@@ -1,0 +1,86 @@
+#### Preamble ####
+# Purpose: Tests the analysis data
+# Author: Kaavya Kalani
+# Date: 6 April 2024
+# Contact: kaavya.kalani@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: Run 01-data_cleaning.R
+
+#### Workspace setup ####
+library(tidyverse)
+
+print("Testing analysis data")
+
+analysis_data <- read.csv("data/analysis_data/expeditions.csv")
+
+# Test 1: Check for missing data
+if (any(is.na(analysis_data))) {
+  print("There are missing data.")
+} else {
+  print("There are no missing data.")
+}
+
+# Test 2: Check the number of columns
+if (ncol(analysis_data) == 9) {
+  print("Number of columns is 9.")
+} else {
+  print("Number of columns is not 9.")
+}
+
+# Test 3: Check the values in the height column are all positive
+if (all(analysis_data$height > 0)) {
+  print("Values in the 'height' column are positive.")
+} else {
+  print("There are invalid values in the 'height' column.")
+}
+
+# Test 4: Check the values in the seasons column are all valid
+seasons <- c("Autumn", "Spring", "Winter", "Summer")
+if (all(analysis_data$seasons %in% seasons)) {
+  print("Values in the 'seasons' column are valid")
+} else {
+  print("There are invalid values in the 'seasons' column.")
+}
+
+# Test 5: Check the values in the sex column are all valid
+if (all(analysis_data$sex %in% c('M', 'F'))) {
+  print("Values in the 'sex' column are valid")
+} else {
+  print("There are invalid values in the 'sex' column.")
+}
+
+# Test 6: Check the values in the age column are all valid
+if (all(analysis_data$age > 0)) {
+  print("Values in the 'age' column are valid")
+} else {
+  print("There are invalid values in the 'age' column.")
+}
+
+# Test 7: Check the values in the success column are all valid
+if (all(analysis_data$success %in% c(TRUE, FALSE))) {
+  print("Values in the 'success' column are valid")
+} else {
+  print("There are invalid values in the 'success' column.")
+}
+
+# Test 8: Check the values in the solo column are all valid
+if (all(analysis_data$solo %in% c(TRUE, FALSE))) {
+  print("Values in the 'solo' column are valid")
+} else {
+  print("There are invalid values in the 'solo' column.")
+}
+
+# Test 9: Check the values in the died column are all valid
[…]

### 2024-04-13 scripts/04-model.R

@@ -1,0 +1,32 @@
+#### Preamble ####
+# Purpose: Bayesian logistic regression model
+# Author: Kaavya Kalani
+# Date: 6 April 2024
+# Contact: kaavya.kalani@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: Run 01-data_cleaning.R
+
+#### Workspace setup ####
+library(tidyverse)
+library(rstanarm)
+
+#### Read the data ####
+analysis_data <- read.csv("data/analysis_data/expeditions.csv")
+
+single_bay <- 
+  stan_glm(
+    success ~ sex + seasons,
+    data = analysis_data,
+    family = binomial(link = "logit"),
+    prior = normal(location = 0, scale = 2.5, autoscale = TRUE),
+    prior_intercept = normal(location = 0, scale = 2.5, autoscale = TRUE),
+    cores = 4,
+    adapt_delta = 0.99,
+    seed = 853
+  )
+
+#### Save model ####
+saveRDS(
+  single_bay,
+  file = "models/single_bay.rds"
+)

### 2024-04-13 paper/paper.qmd
Model description completed

@@
 # Model {#sec-model}
-I used a Bayesian Logistic Regression model to do XXX. Logistic regression is a method used for binary classification to predict the probability of a categorical dependent variable.
+I used a Bayesian Logistic Regression model to find the probability that someone will successfully complete the Himalayan peak they are on the expedition for.
+Logistic regression is a method used for binary classification to predict the probability of a categorical dependent variable.
 
-For my analysis, a logistic regression model will be first used to model XXX. The model will be based on XXX independent demographic variables: XXX. (Use backticks)
+For my analysis, a logistic regression model will be used to model if the person will be successful on their attempt to climb the particular Himalayan peak.
+The model will be based on five independent demographic variables: `height of the peak`, `sex`, `age`, `seasons` and `solo`.
 

### 2024-04-13 paper/paper.qmd
Data section updated

@@
 title: "Himalayan Expeditions Success Analysis "
-subtitle: "Analysis using data from XXX leveraging Bayesian Logistic Regression"
+subtitle: "Analysis using data from Himalayan expeditions from 1905 through Spring 2019 leveraging Bayesian Logistic Regression"
 author: Kaavya Kalani
@@
 ## Analysis Dataset
-The analysis data is from XXXX. This is how it was collected.
+The raw datasets were obtained from Alex Cookson's datasets (link). I chose the ones cleaned for Himalayan expeditions. Alex got his datasets from The Himalayan Database (link).
 
-A person becomes an entry in this dataset if XXX. (Measurement)
+The Himalayan Database is a compilation of records for all expeditions that have climbed in the Nepal Himalaya.
+The database is based on the expedition archives of Elizabeth Hawley, a longtime journalist based in Kathmandu, and it is supplemented by information gathered from books, alpine journals and correspondence with Himalayan climbers.
 
-The data from datasets for XXX are combined to form our anaysis dataset. Among the overall range of variables available, we chose XXX to be included in our analysis dataset.
+The original database currently covers all expeditions from 1905 through Spring-Summer 2023 to the most significant mountaineering peaks in Nepal.
+Also included are expeditions to both sides of border peaks such as Everest, Cho Oyu, Makalu and Kangchenjunga as well as to some smaller border peaks.
+Data on expeditions to trekking peaks are included for early attempts, first ascents and major accidents.
+The updates to this database are published bi-annually.
 
-- Describe the variables
+My dataset derived from Alex's contains the entries from 1905 through Spring 2019.
 
-Out of these, we will be using XXX in our model as the independent variables and XXX as the dependent variable.
+The three datasets I considered included information about all peaks, all expeditions on those peaks and all members on those expeditions.
+The data from these three datasets are combined to form our analysis dataset.
+
+A person becomes an entry in my analysis dataset if, between 1905 and Spring 2019, they attempted to climb any one of the many Himalayan peaks in Nepal.
+It also included expeditions to both sides of border peaks as mentioned before.
+
+Among the overall range of variables available, we chose the following to be included in our analysis dataset.
+
+- `Height of the peak` in metres for the peak the person in the current entry is on an expedition for
+- `Seasons` is the season the expedition is in. This takes on either of the 4 values: Autumn, Spring, Winter, Summer.
+- `Sex` is the sex reported by the expedition member and it is either male or female.
+- `Age` is the age of the expedition member at the time of the expedition.
+Depending on the best available data, this could be as of the summit date, the date of death, or the date of arrival at basecamp.
+- `Success`represents whether the person was successful in summitting the goal peak.
+- `Solo` represents whether the person attempted a solo ascent.
+- `Died` represents whether the person died during the expedition.
+
+Out of these, we will be using `height of the peak`, `seasons`, `sex`, `age` and `solo` in our model as the independent variables and `success` as the dependent variable.
 

### 2024-04-13 scripts/02-data_cleaning.R

@@ -16,7 +16,7 @@ members <- read.csv("data/raw_data/members.csv")
 
 merged_data1 <- merge(expeditions, members, by = "expedition_id")
 merged_data1 <- merged_data1 %>%
-  select(peak_id.x, season.x, sex, age, success, solo, died, members) %>%
+  select(peak_id.x, season.x, sex, age, success, solo, died) %>%
   rename(peak_id = peak_id.x)
 
 peaks <- peaks %>%
@@ -24,8 +24,8 @@ peaks <- peaks %>%
 
 merged_data2 <- merge(merged_data1, peaks, by = "peak_id")
 main_data <- merged_data2 %>%
-  filter(!is.na(sex), !is.na(age), members > 0) %>%
-  select(peak_id, height_metres, season.x, sex, age, success, solo, died, members) %>%
+  filter(!is.na(sex), !is.na(age)) %>%
+  select(peak_id, height_metres, season.x, sex, age, success, solo, died) %>%
   rename(height = height_metres, seasons = season.x)
 
 #### Save the cleaned data ####

### 2024-04-13 scripts/03-test_data.R

@@ -21,10 +21,10 @@ if (any(is.na(analysis_data))) {
 }
 
 # Test 2: Check the number of columns
-if (ncol(analysis_data) == 9) {
-  print("Number of columns is 9.")
+if (ncol(analysis_data) == 8) {
+  print("Number of columns is 8.")
 } else {
-  print("Number of columns is not 9.")
+  print("Number of columns is not 8.")
 }
 
 # Test 3: Check the values in the height column are all positive
@@ -77,10 +77,4 @@ if (all(analysis_data$died %in% c(TRUE, FALSE))) {
   print("There are invalid values in the 'died' column.")
 }
 
-# Test 10: Check the values in the members column are all valid
-if (all(analysis_data$members > 0)) {
-  print("Values in the 'members' column are valid")
-} else {
-  print("There are invalid values in the 'members' column.")
-}
 

### 2024-04-13 paper/paper.qmd
Cleaning information added to the appendix

@@
 - `Died` represents whether the person died during the expedition.
+
+(Plot the variables)
 
@@
 
+## Cleaning
+For the analysis data, the cleaning steps we took were:
+
+1. Initial merging: I first merge the raw data from the `expeditions` and `members` datasets based on a common identifier, `expedition_id`, to consolidate information about expedition participants.
+
+2.
+Column selection and renaming: Irrelevant columns are removed from the merged dataset, and the remaining columns (`peak_id.x`, `season.x`, `sex`, `age`, `success`, `solo`, `died`) are selected for further analysis.
+Additionally, column `peak_id.x` is renamed to `peak_id`.
+
+3.
+Secondary merging: The cleaned `expeditions` dataset is merged with the `peaks` dataset based on a common identifier, `peak_id`, to incorporate information about the height of each peak climbed during expeditions.
+
+4. Filtering out incomplete data: Rows with missing values for key variables such as `sex` or `age` are filtered out to ensure the integrity of the dataset.
+
+5.
+Final dataset creation: The resulting dataset is further refined to include only relevant columns (`peak_id`, `height_metres`, `season.x`, `sex`, `age`, `success`, `solo`, `died`).
+Some column names are adjusted for clarity, such as `season.x` being renamed to `seasons`, and `height_metres` being renamed to `height`.
+
 ## Analysis dataset

### 2024-04-14 paper/paper.qmd
Introduction added

@@
 # Introduction
+Mountaineering, with its blend of adventure and challenge, has captivated explorers for generations.
+It represents a pinnacle of human endeavor, testing physical prowess, mental fortitude, and strategic planning against some of the most formidable natural landscapes on Earth.
+Amidst the allure of conquering these majestic peaks, lies a critical question: what factors contribute to the likelihood of a successful summit attempt?
+This paper delves into precisely this inquiry.
 
-1) broader context to motivate; 
-2) some detail about what the paper is about; 
-3) a clear gap that needs to be filled; 
-4) what was done; 
-5) what was found; 
-6) why it is important;
-7) Estimand
+There are numerous mountain ranges around the world where people embark on expeditions.
+One of the most famous mountain ranges are the Himalayas, which also consist of the highest mountain peak in the world, Mt.
+Everest.
+Our paper uses the data from expeditions to the Himalayan mountain ranges in Nepal.
 
-(Likely 3 or 4 paragraphs, or 10 per cent of total.)
+Our estimand is the relationship between different demographic, environmental and geographic factors (such as height of the peak, sex, age, season of expedition and if it was a solo ascent) and the successfulness of climbing a summit.
+Using our extensive analysis dataset, our goal is to identify trends and factors that influence a successful expedition and eventually conclude what factors help in a more successful attempt.
+
+We use data from Alex Cookson’s datasets (link) which are sourced from The Himalayan Expedition records (link), to understand these factors and trends.
+This is done by leveraging a Bayesian Logistic Regression model and then predicting the probability of a successful attempt over various demographic and environmental factors.
+
+Our analysis led to us … (add what was found)
+
+This research highlights the interplay between climber attributes and environmental conditions, offering insights into Himalayan expeditions.
+These findings have practical implications for expedition planning, risk management, and safety protocols.
+Ultimately, our study aims to improve decision-making in high-altitude mountaineering, making it safer and more informed in one of the world's most challenging environments.
 
@@
 # Model {#sec-model}
-I used a Bayesian Logistic Regression model to find the probability that someone will successfully complete the Himalayan peak they are on the expedition for.
+I used a Bayesian Logistic Regression model to find the probability that someone will successfully summit the Himalayan peak they are on the expedition for.
 Logistic regression is a method used for binary classification to predict the probability of a categorical dependent variable.
 
-For my analysis, a logistic regression model will be used to model if the person will be successful on their attempt to climb the particular Himalayan peak.
+For my analysis, a logistic regression model will be used to model if the person will be successful on their attempt to summit the particular Himalayan peak.
 The model will be based on five independent demographic variables: `height of the peak`, `sex`, `age`, `seasons` and `solo`.
@@
 - $\beta_1$ is the coefficient corresponding to height of the peak
-- $\beta_2$ is the coefficient corresponding to sex
-- $\beta_3$ is the coefficients corresponding to age
-- $\beta_4$ is the coefficients corresponding to seasons
-- $\beta_5$ is the coefficients corresponding to solo
+- $\beta_2$ is the coefficient corresponding to sex of the person
+- $\beta_3$ is the coefficients corresponding to age of the person
+- $\beta_4$ is the coefficients corresponding to seasons of the person
+- $\beta_5$ is the coefficients corresponding to if the person is attempting the summit alone.
 
@@
 
-# Create a data frame with all combinations of sex and seasons
-sex_levels <- c('M', 'F')  # Replace with actual categories
-seasons_levels <- unique(analysis_data$seasons)  # Extract unique seasons from your analysis data
-
-new_data <- expand.grid(
-  sex = sex_levels,
-  seasons = seasons_levels
-)
-
-# Predicting success based on sex and seasons
-predictions <- predict(expeditions_model, newdata = new_data, type = "response")
-
-# Create a data frame to display the predictions
-prediction_df <- data.frame(
-  sex = new_data$sex, 
-  seasons = new_data$seasons, 
-  success_probability = predictions
-)
-
 ggplot(prediction_df, aes(x = sex, y = success_probability, fill = sex)) +

### 2024-04-14 paper/paper.qmd
Updates to multiple files (old versions so that the qmd renders as current model is not working properly)

@@
 @sec-results then highlights and discusses the trends and associations found during the analysis.
-Lastly,@sec-discussion talks about some interesting trends found in @sec-results in depth, link it to the real world and also highlight the weaknesses and future of my analysis.
+Lastly, @sec-discussion talks about some interesting trends found in @sec-results in depth, link it to the real world and also highlight the weaknesses and future of my analysis.
 
@@
 - $\beta_4$ is the coefficients corresponding to seasons of the person
-- $\beta_5$ is the coefficients corresponding to if the person is attempting the summit alone.
+- $\beta_5$ is the coefficients corresponding to if the person is attempting the summit alone
 
@@
 
+# Create a data frame with all combinations of sex and seasons
+sex_levels <- unique(analysis_data$sex) 
+seasons_levels <- unique(analysis_data$seasons)
+
+new_data <- expand.grid(
+  sex = sex_levels,
+  seasons = seasons_levels
+)
+
+# Predicting success based on sex and seasons
+predictions <- predict(expeditions_model, newdata = new_data, type = "response")
+
+# Create a data frame to display the predictions
+prediction_df <- data.frame(
+  sex = new_data$sex, 
+  seasons = new_data$seasons, 
+  success_probability = predictions
+)
+
 ggplot(prediction_df, aes(x = sex, y = success_probability, fill = sex)) +
@@
 2) What is something that we learn about the world?
-3) What are some weaknesses of what was done? What is left to learn or how should we proceed in the future?
 

### 2024-04-14 scripts/02-data_cleaning.R

@@ -1,7 +1,7 @@
 #### Preamble ####
 # Purpose: Cleans the raw data
 # Author: Kaavya Kalani
-# Date:6 April 2024
+# Date: 6 April 2024
 # Contact: kaavya.kalani@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: None
@@ -25,8 +25,21 @@ peaks <- peaks %>%
 merged_data2 <- merge(merged_data1, peaks, by = "peak_id")
 main_data <- merged_data2 %>%
   filter(!is.na(sex), !is.na(age)) %>%
-  select(peak_id, height_metres, season.x, sex, age, success, solo, died) %>%
-  rename(height = height_metres, seasons = season.x)
+  rename(height = height_metres, seasons = season.x) %>%
+  mutate(height_range = case_when(
+    height >= 5400 & height < 5750 ~ "5400 - 5749",
+    height >= 5750 & height < 6100 ~ "5750 - 6099",
+    height >= 6100 & height < 6450 ~ "6100 - 6449",
+    height >= 6450 & height < 6800 ~ "6450 - 6799",
+    height >= 6800 & height < 7150 ~ "6800 - 7149",
+    height >= 7150 & height < 7500 ~ "7150 - 7499",
+    height >= 7500 & height < 7850 ~ "7500 - 7849",
+    height >= 7850 & height < 8200 ~ "7850 - 8199",
+    height >= 8200 & height < 8550 ~ "8200 - 8549",
+    height >= 8550 & height <= 8900 ~ "8550 - 8900",
+    TRUE ~ as.character(height)
+    )) %>%
+    select(peak_id, height_range, seasons, sex, age, success, solo, died)
 
 #### Save the cleaned data ####
 write.csv(main_data, "data/analysis_data/expeditions.csv", row.names = FALSE)

### 2024-04-14 scripts/04-model.R

@@ -4,7 +4,7 @@
 # Date: 6 April 2024
 # Contact: kaavya.kalani@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: Run 01-data_cleaning.R
+# Pre-requisites: Run 02-data_cleaning.R
 
 #### Workspace setup ####
 library(tidyverse)

### 2024-04-14 paper/paper.qmd
Abstract added

@@
 ---
-title: "Himalayan Expeditions Success Analysis "
+title: "Himalayan Expeditions Success Analysis"
 subtitle: "Analysis using data from Himalayan expeditions from 1905 through Spring 2019 leveraging Bayesian Logistic Regression"
 author: Kaavya Kalani
-thanks: "Code and data supporting this analysis is available at: https://github.com/kaavyakalani26/himalayan-expeditions-analysis"
+thanks: "Code and data supporting this analysis is available at: [https://github.com/kaavyakalani26/himalayan-expeditions-analysis](https://github.com/kaavyakalani26/himalayan-expeditions-analysis)"
 date: today
 date-format: long
-abstract: "Abstract"
+abstract: "I analyse the relationship between different demographic, environmental and geographic factors and the successfulness of an attempt to summit a peak.
+Data from Himalayan expeditions in Nepal from 1905 through Spring 2019 is used and a Bayesian Logistic Regression model is leveraged to analyse the trends and factors influencing a successful summit.
+We find a strong relations between height of the peak, sex, age, season of expedition, if it was a solo ascent, and the success in summitting.
+This research provides insights into factors affecting the successfullness of an expedition to a summit and helps for future expedition planning, risk management, and safety protocols."
 format: pdf

### 2024-04-18 scripts/02-data_cleaning.R
Final model created and predictions added

@@ -38,8 +38,19 @@ main_data <- merged_data2 %>%
     height >= 8200 & height < 8550 ~ "8200 - 8549",
     height >= 8550 & height <= 8900 ~ "8550 - 8900",
     TRUE ~ as.character(height)
+    ),
+    age_range = case_when(
+      age <= 18 ~ "Under 18",
+      age > 18 & age <= 30 ~ "19-30",
+      age > 30 & age <= 40 ~ "31-40",
+      age > 40 & age <= 50 ~ "41-50",
+      age > 50 & age <= 60 ~ "51-60",
+      age > 60 & age <= 70 ~ "61-70",
+      age > 70 & age <= 80 ~ "71-80",
+      age > 80 & age <= 90 ~ "81-90",
+      TRUE ~ as.character(age)
     )) %>%
-    select(peak_id, height_range, seasons, sex, age, success, solo, died)
+    select(peak_id, height_range, seasons, sex, age, age_range, success, solo, died)
 
 #### Save the cleaned data ####
 write.csv(main_data, "data/analysis_data/expeditions.csv", row.names = FALSE)

### 2024-04-18 scripts/04-model.R

@@ -14,8 +14,8 @@ library(rstanarm)
 analysis_data <- read.csv("data/analysis_data/expeditions.csv")
 
 single_bay <- 
-  stan_glm(
-    success ~ sex + seasons,
+  stan_glm(  
+    success ~ height_range + sex + solo + age_range + seasons,
     data = analysis_data,
     family = binomial(link = "logit"),
     prior = normal(location = 0, scale = 2.5, autoscale = TRUE),

### 2024-04-18 scripts/05-predictions.R

@@ -1,0 +1,46 @@
+#### Preamble ####
+# Purpose: Creates prediction dataset to use in Results of the paper
+# Author: Kaavya Kalani
+# Date: 13 April 2024
+# Contact: kaavya.kalani@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: Run 04-model.R
+
+#### Workspace setup ####
+library(tidyverse)
+
+#### Prediciton dataset ####
+analysis_data <- read.csv("data/analysis_data/expeditions.csv")
+expeditions_model <- readRDS("models/single_bay.rds")
+
+# Create a data frame with all combinations of variables
+sex_levels <- unique(analysis_data$sex) 
+seasons_levels <- unique(analysis_data$seasons)
+height_levels <- unique(analysis_data$height_range)
+age_levels <- unique(analysis_data$age_range)
+solo_levels <- unique(analysis_data$solo)
+
+new_data <- expand.grid(
+  sex = sex_levels,
+  seasons = seasons_levels,
+  height_range = height_levels,
+  age_range = age_levels,
+  solo = solo_levels
+)
+
+# Predicting success based on all variables
+predictions <- predict(expeditions_model, newdata = new_data, type = "response")
+
+# Create a data frame to display the predictions
+prediction_df <- data.frame(
+  sex = new_data$sex, 
+  seasons = new_data$seasons, 
+  height_range = new_data$height_range,
+  age_range = new_data$age_range,
+  solo = new_data$solo,
+  success_probability = predictions
+)
+
+#### Saving the predictions ####
+write.csv(prediction_df, "data/predictions/predictions.csv", row.names = FALSE)
+

### 2024-04-18 paper/paper.qmd
Analysis data stored as a parquet

@@
 library(kableExtra)
+library(arrow)
 
 expeditions_model <- readRDS(here("models/single_bay.rds"))
-analysis_data <- read.csv(here("data/analysis_data/expeditions.csv"))
+analysis_data <- read_parquet("data/analysis_data/expeditions.parquet")
+predictions <- read.csv(here("data/predictions/predictions.csv"))
 ```
@@
 For this analysis, we have used combined three datasets into one, which is used for analysis.
-The datasets were cleaned and analysed using the statistical programming software `R` [@citeR] along with the help `tidyverse` [@citeTidyverse], `knitr` [@citeKnitr], `ggplot2` [@citeGgplot], `here` [@citeHere], `dplyr` [@citeDplyr], `rstanarm` [@citeRstanarm], `broom.mixed` [@citeBroomMixed], `modelsummary` [@citeModelSummary] and `kableExtra` [@citeKableExtra].
+The datasets were cleaned and analysed using the statistical programming software `R` [@citeR] along with the help `tidyverse` [@citeTidyverse], `knitr` [@citeKnitr], `ggplot2` [@citeGgplot], `here` [@citeHere], `dplyr` [@citeDplyr], `rstanarm` [@citeRstanarm], `arrow` [@citeArrow], `broom.mixed` [@citeBroomMixed], `modelsummary` [@citeModelSummary] and `kableExtra` [@citeKableExtra].
 
@@
 #| tbl-cap: Summary of the model
-
-# Create model summary
+# 
+# # Create model summary
 # model_summary <- modelsummary(expeditions_model)
@@
 #| fig-cap: Coefficients of the model
+
+coefficients <- coefficients[coefficients$term != "age_range81-90", ]
 
@@
 
-# Create a data frame with all combinations of sex and seasons
-sex_levels <- unique(analysis_data$sex) 
-seasons_levels <- unique(analysis_data$seasons)
-
-new_data <- expand.grid(
-  sex = sex_levels,
-  seasons = seasons_levels
-)
-
-# Predicting success based on sex and seasons
-predictions <- predict(expeditions_model, newdata = new_data, type = "response")
-
-# Create a data frame to display the predictions
-prediction_df <- data.frame(
-  sex = new_data$sex, 
-  seasons = new_data$seasons, 
-  success_probability = predictions
-)
-
-ggplot(prediction_df, aes(x = sex, y = success_probability, fill = sex)) +
+ggplot(predictions, aes(x = sex, y = success_probability, fill = sex)) +
   geom_bar(stat = "identity", position = "dodge") +

### 2024-04-18 scripts/00-install-packages.R

@@ -15,7 +15,8 @@ packages <- c("tidyverse",
               "ggplot2", 
               "modelsummary", 
               "here", 
-              "kableExtra")
+              "kableExtra",
+              "arrow")
 
 # Install packages if they are not already installed
 for (package in packages) {

### 2024-04-18 scripts/02-data_cleaning.R

@@ -8,6 +8,7 @@
 
 #### Workspace setup ####
 library(tidyverse)
+library(arrow)
 
 #### Load and clean data ####
 peaks <- read.csv("data/raw_data/peaks.csv")
@@ -26,18 +27,19 @@ merged_data2 <- merge(merged_data1, peaks, by = "peak_id")
 main_data <- merged_data2 %>%
   filter(!is.na(sex), !is.na(age)) %>%
   rename(height = height_metres, seasons = season.x) %>%
-  mutate(height_range = case_when(
-    height >= 5400 & height < 5750 ~ "5400 - 5749",
-    height >= 5750 & height < 6100 ~ "5750 - 6099",
-    height >= 6100 & height < 6450 ~ "6100 - 6449",
-    height >= 6450 & height < 6800 ~ "6450 - 6799",
-    height >= 6800 & height < 7150 ~ "6800 - 7149",
-    height >= 7150 & height < 7500 ~ "7150 - 7499",
-    height >= 7500 & height < 7850 ~ "7500 - 7849",
-    height >= 7850 & height < 8200 ~ "7850 - 8199",
-    height >= 8200 & height < 8550 ~ "8200 - 8549",
-    height >= 8550 & height <= 8900 ~ "8550 - 8900",
-    TRUE ~ as.character(height)
+  mutate(
+    height_range = case_when(
+      height >= 5400 & height < 5750 ~ "5400 - 5749",
+      height >= 5750 & height < 6100 ~ "5750 - 6099",
+      height >= 6100 & height < 6450 ~ "6100 - 6449",
+      height >= 6450 & height < 6800 ~ "6450 - 6799",
+      height >= 6800 & height < 7150 ~ "6800 - 7149",
+      height >= 7150 & height < 7500 ~ "7150 - 7499",
+      height >= 7500 & height < 7850 ~ "7500 - 7849",
+      height >= 7850 & height < 8200 ~ "7850 - 8199",
+      height >= 8200 & height < 8550 ~ "8200 - 8549",
+      height >= 8550 & height <= 8900 ~ "8550 - 8900",
+      TRUE ~ as.character(height)
     ),
     age_range = case_when(
       age <= 18 ~ "Under 18",
@@ -49,8 +51,9 @@ main_data <- merged_data2 %>%
       age > 70 & age <= 80 ~ "71-80",
       age > 80 & age <= 90 ~ "81-90",
       TRUE ~ as.character(age)
-    )) %>%
[…]

### 2024-04-18 scripts/03-test_data.R

@@ -8,10 +8,12 @@
 
 #### Workspace setup ####
 library(tidyverse)
+library(arrow)
 
+#### Tests ####
 print("Testing analysis data")
 
-analysis_data <- read.csv("data/analysis_data/expeditions.csv")
+analysis_data <- read_parquet("data/analysis_data/expeditions.parquet")
 
 # Test 1: Check for missing data
 if (any(is.na(analysis_data))) {

### 2024-04-18 scripts/04-model.R

@@ -9,9 +9,10 @@
 #### Workspace setup ####
 library(tidyverse)
 library(rstanarm)
+library(arrow)
 
 #### Read the data ####
-analysis_data <- read.csv("data/analysis_data/expeditions.csv")
+analysis_data <- read_parquet("data/analysis_data/expeditions.parquet")
 
 single_bay <- 
   stan_glm(  

### 2024-04-18 scripts/05-predictions.R

@@ -8,9 +8,10 @@
 
 #### Workspace setup ####
 library(tidyverse)
+library(arrow)
 
 #### Prediciton dataset ####
-analysis_data <- read.csv("data/analysis_data/expeditions.csv")
+analysis_data <- read_parquet("data/analysis_data/expeditions.parquet")
 expeditions_model <- readRDS("models/single_bay.rds")
 
 # Create a data frame with all combinations of variables

### 2024-04-18 paper/paper.qmd
Simulation added and comments to all scripts added

@@
 expeditions_model <- readRDS(here("models/single_bay.rds"))
-analysis_data <- read_parquet("data/analysis_data/expeditions.parquet")
+analysis_data <- read_parquet(here("data/analysis_data/expeditions.parquet"))
 predictions <- read.csv(here("data/predictions/predictions.csv"))

### 2024-04-18 scripts/00-install-packages.R

@@ -1,7 +1,7 @@
 #### Preamble ####
 # Purpose: Install all necessary packages for this paper's replication
 # Author: Kaavya Kalani
-# Date:6 April 2024
+# Date: 18 April 2024
 # Contact: kaavya.kalani@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: None

### 2024-04-18 scripts/01-simulate_data.R

@@ -1,9 +1,41 @@
 #### Preamble ####
-# Purpose: Simulated the survey and poststratification datasets
+# Purpose: Simulated the analysis dataset
 # Author: Kaavya Kalani
-# Date: XX March 2024
+# Date: 18 March 2024
 # Contact: kaavya.kalani@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: None
 
 #### Workspace setup ####
+library(tidyverse)
+
+#### Simulation ####
+# Set seed for reproducibility
+set.seed(823)
+
+# Number of rows in the dataset
+n <- 1000
+
+# Generate age values
+age <- round(runif(n, min = 15, max = 90))
+
+# Generate sex values
+sex <- sample(c("m", "f"), n, replace = TRUE)
+
+# Generate season values
+season <- sample(c("summer", "winter", "spring", "autumn"), n, replace = TRUE)
+
+# Generate peak height values
+peak_height <- round(runif(n, min = 5000, max = 8849))
+
+# Generate success values
+success <- sample(0:1, n, replace = TRUE)
+
+# Generate died values based on success
+died <- ifelse(success == 1, rbinom(n, 1, 0.1), rbinom(n, 1, 0.5))
+
+# Create dataframe
+data <- data.frame(age, sex, season, peak_height, success, died)
+
+# View the first few rows of the dataset
+head(data)

### 2024-04-18 scripts/02-data_cleaning.R

@@ -1,7 +1,7 @@
 #### Preamble ####
 # Purpose: Cleans the raw data
 # Author: Kaavya Kalani
-# Date: 6 April 2024
+# Date: 18 April 2024
 # Contact: kaavya.kalani@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: None
@@ -11,23 +11,31 @@ library(tidyverse)
 library(arrow)
 
 #### Load and clean data ####
+# Read the raw data files
 peaks <- read.csv("data/raw_data/peaks.csv")
 expeditions <- read.csv("data/raw_data/expeditions.csv")
 members <- read.csv("data/raw_data/members.csv")
 
+# Merge expedition and member data based on expedition_id
 merged_data1 <- merge(expeditions, members, by = "expedition_id")
+
+# Select relevant columns and rename peak_id
 merged_data1 <- merged_data1 %>%
   select(peak_id.x, season.x, sex, age, success, solo, died) %>%
   rename(peak_id = peak_id.x)
 
+# Select relevant columns from the peaks data
 peaks <- peaks %>%
   select(peak_id, height_metres)
 
+# Merge the previously merged data with peaks data based on peak_id
 merged_data2 <- merge(merged_data1, peaks, by = "peak_id")
+
 main_data <- merged_data2 %>%
-  filter(!is.na(sex), !is.na(age)) %>%
+  filter(!is.na(sex), !is.na(age)) %>% # Filter out rows with missing values for sex or age
   rename(height = height_metres, seasons = season.x) %>%
   mutate(
+    # Create height range categories
     height_range = case_when(
       height >= 5400 & height < 5750 ~ "5400 - 5749",
       height >= 5750 & height < 6100 ~ "5750 - 6099",
@@ -41,6 +49,7 @@ main_data <- merged_data2 %>%
       height >= 8550 & height <= 8900 ~ "8550 - 8900",
       TRUE ~ as.character(height)
     ),
+    # Create age range categories
     age_range = case_when(
       age <= 18 ~ "Under 18",
       age > 18 & age <= 30 ~ "19-30",
@@ -53,7 +62,8 @@ main_data <- merged_data2 %>%
       TRUE ~ as.character(age)
     )
   ) %>%
-  select(peak_id, height_range, seasons, sex, age, age_range, success, solo, died)
+  # Select final columns
+  select(peak_id, height_range, seasons, sex, age_range, success, solo, died)
 
 #### Save the cleaned data ####
[…]

### 2024-04-18 scripts/03-test_data.R

@@ -1,7 +1,7 @@
 #### Preamble ####
 # Purpose: Tests the analysis data
 # Author: Kaavya Kalani
-# Date: 6 April 2024
+# Date: 18 April 2024
 # Contact: kaavya.kalani@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: Run 02-data_cleaning.R

### 2024-04-18 scripts/04-model.R

@@ -1,7 +1,7 @@
 #### Preamble ####
 # Purpose: Bayesian logistic regression model
 # Author: Kaavya Kalani
-# Date: 6 April 2024
+# Date: 18 April 2024
 # Contact: kaavya.kalani@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: Run 02-data_cleaning.R
@@ -11,9 +11,11 @@ library(tidyverse)
 library(rstanarm)
 library(arrow)
 
-#### Read the data ####
+#### Read the data and create model ####
+# Read the cleaned analysis dataset
 analysis_data <- read_parquet("data/analysis_data/expeditions.parquet")
 
+# Fit a Bayesian logistic regression model
 single_bay <- 
   stan_glm(  
     success ~ height_range + sex + solo + age_range + seasons,

### 2024-04-18 scripts/02-data_cleaning.R
Tests updated

@@ -63,7 +63,7 @@ main_data <- merged_data2 %>%
     )
   ) %>%
   # Select final columns
-  select(peak_id, height_range, seasons, sex, age_range, success, solo, died)
+  select(height_range, seasons, sex, age_range, success, solo, died)
 
 #### Save the cleaned data ####
 write_parquet(main_data, "data/analysis_data/expeditions.parquet")

### 2024-04-18 scripts/03-test_data.R

@@ -23,15 +23,17 @@ if (any(is.na(analysis_data))) {
 }
 
 # Test 2: Check the number of columns
-if (ncol(analysis_data) == 8) {
-  print("Number of columns is 8.")
+if (ncol(analysis_data) == 7) {
+  print("Number of columns is 7.")
 } else {
-  print("Number of columns is not 8.")
+  print("Number of columns is not 7.")
 }
 
-# Test 3: Check the values in the height column are all positive
-if (all(analysis_data$height > 0)) {
-  print("Values in the 'height' column are positive.")
+# Test 3: Check the values in the height column are all valid
+heights <- c("5400 - 5749", "5750 - 6099", "6100 - 6449", "6450 - 6799", "6800 - 7149", "7150 - 7499", 
+             "7500 - 7849", "7850 - 8199", "8200 - 8549", "8550 - 8900")
+if (all(analysis_data$height_range %in% heights)) {
+  print("Values in the 'height' column are valid")
 } else {
   print("There are invalid values in the 'height' column.")
 }
@@ -52,7 +54,8 @@ if (all(analysis_data$sex %in% c('M', 'F'))) {
 }
 
 # Test 6: Check the values in the age column are all valid
-if (all(analysis_data$age > 0)) {
+ages <- c("Under 18", "19-30", "31-40", "41-50", "51-60", "61-70", "71-80", "81-90")
+if (all(analysis_data$age_range %in% ages)) {
   print("Values in the 'age' column are valid")
 } else {
   print("There are invalid values in the 'age' column.")

### 2024-04-18 scripts/05-predictions.R

@@ -1,7 +1,7 @@
 #### Preamble ####
 # Purpose: Creates prediction dataset to use in Results of the paper
 # Author: Kaavya Kalani
-# Date: 13 April 2024
+# Date: 18 April 2024
 # Contact: kaavya.kalani@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: Run 04-model.R

### 2024-04-18 scripts/03-test_data.R
A boolean checked added to provide test summary at the end of the script being run

@@ -15,71 +15,88 @@ print("Testing analysis data")
 
 analysis_data <- read_parquet("data/analysis_data/expeditions.parquet")
 
+# Initialize a boolean variable to track test results
+all_tests_passed <- TRUE
+
 # Test 1: Check for missing data
 if (any(is.na(analysis_data))) {
   print("There are missing data.")
+  all_tests_passed <- FALSE
 } else {
   print("There are no missing data.")
 }
 
 # Test 2: Check the number of columns
-if (ncol(analysis_data) == 7) {
-  print("Number of columns is 7.")
-} else {
+if (ncol(analysis_data) != 7) {
   print("Number of columns is not 7.")
+  all_tests_passed <- FALSE
+} else {
+  print("Number of columns is 7.")
 }
 
 # Test 3: Check the values in the height column are all valid
 heights <- c("5400 - 5749", "5750 - 6099", "6100 - 6449", "6450 - 6799", "6800 - 7149", "7150 - 7499", 
              "7500 - 7849", "7850 - 8199", "8200 - 8549", "8550 - 8900")
-if (all(analysis_data$height_range %in% heights)) {
-  print("Values in the 'height' column are valid")
-} else {
+if (!all(analysis_data$height_range %in% heights)) {
   print("There are invalid values in the 'height' column.")
+  all_tests_passed <- FALSE
+} else {
+  print("Values in the 'height' column are valid")
 }
 
 # Test 4: Check the values in the seasons column are all valid
 seasons <- c("Autumn", "Spring", "Winter", "Summer")
-if (all(analysis_data$seasons %in% seasons)) {
-  print("Values in the 'seasons' column are valid")
-} else {
+if (!all(analysis_data$seasons %in% seasons)) {
   print("There are invalid values in the 'seasons' column.")
+  all_tests_passed <- FALSE
+} else {
+  print("Values in the 'seasons' column are valid")
 }
 
 # Test 5: Check the values in the sex column are all valid
-if (all(analysis_data$sex %in% c('M', 'F'))) {
-  print("Values in the 'sex' column are valid")
-} else {
+if (!all(analysis_data$sex %in% c('M', 'F'))) {
   print("There are invalid values in the 'sex' column.")
+  all_tests_passed <- FALSE
+} else {
+  print("Values in the 'sex' column are valid")
 }
 
 # Test 6: Check the values in the age column are all valid
 ages <- c("Under 18", "19-30", "31-40", "41-50", "51-60", "61-70", "71-80", "81-90")
[…]

### 2024-04-18 scripts/00-install-packages.R
styler used on all scripts

@@ -7,16 +7,18 @@
 # Pre-requisites: None
 
 # List of packages to install
-packages <- c("tidyverse", 
-              "rstanarm", 
-              "broom.mixed", 
-              "dplyr", 
-              "knitr", 
-              "ggplot2", 
-              "modelsummary", 
-              "here", 
-              "kableExtra",
-              "arrow")
+packages <- c(
+  "tidyverse",
+  "rstanarm",
+  "broom.mixed",
+  "dplyr",
+  "knitr",
+  "ggplot2",
+  "modelsummary",
+  "here",
+  "kableExtra",
+  "arrow"
+)
 
 # Install packages if they are not already installed
 for (package in packages) {
@@ -24,4 +26,3 @@ for (package in packages) {
     install.packages(package, dependencies = TRUE)
   }
 }
-

### 2024-04-18 scripts/03-test_data.R

@@ -35,8 +35,10 @@ if (ncol(analysis_data) != 7) {
 }
 
 # Test 3: Check the values in the height column are all valid
-heights <- c("5400 - 5749", "5750 - 6099", "6100 - 6449", "6450 - 6799", "6800 - 7149", "7150 - 7499", 
-             "7500 - 7849", "7850 - 8199", "8200 - 8549", "8550 - 8900")
+heights <- c(
+  "5400 - 5749", "5750 - 6099", "6100 - 6449", "6450 - 6799", "6800 - 7149", "7150 - 7499",
+  "7500 - 7849", "7850 - 8199", "8200 - 8549", "8550 - 8900"
+)
 if (!all(analysis_data$height_range %in% heights)) {
   print("There are invalid values in the 'height' column.")
   all_tests_passed <- FALSE
@@ -54,7 +56,7 @@ if (!all(analysis_data$seasons %in% seasons)) {
 }
 
 # Test 5: Check the values in the sex column are all valid
-if (!all(analysis_data$sex %in% c('M', 'F'))) {
+if (!all(analysis_data$sex %in% c("M", "F"))) {
   print("There are invalid values in the 'sex' column.")
   all_tests_passed <- FALSE
 } else {
@@ -95,8 +97,8 @@ if (!all(analysis_data$died %in% c(TRUE, FALSE))) {
 }
 
 # Print summary based on test results
-if(all_tests_passed) {
+if (all_tests_passed) {
   print("All tests passed.")
 } else {
-  print("One or more tests failed. Check the individual test results for more information on which tests failed.")
+  print("One or more tests failed.")
 }

### 2024-04-18 scripts/04-model.R

@@ -16,8 +16,8 @@ library(arrow)
 analysis_data <- read_parquet("data/analysis_data/expeditions.parquet")
 
 # Fit a Bayesian logistic regression model
-single_bay <- 
-  stan_glm(  
+single_bay <-
+  stan_glm(
     success ~ height_range + sex + solo + age_range + seasons,
     data = analysis_data,
     family = binomial(link = "logit"),

### 2024-04-18 scripts/05-predictions.R

@@ -15,7 +15,7 @@ analysis_data <- read_parquet("data/analysis_data/expeditions.parquet")
 expeditions_model <- readRDS("models/single_bay.rds")
 
 # Create a data frame with all combinations of variables
-sex_levels <- unique(analysis_data$sex) 
+sex_levels <- unique(analysis_data$sex)
 seasons_levels <- unique(analysis_data$seasons)
 height_levels <- unique(analysis_data$height_range)
 age_levels <- unique(analysis_data$age_range)
@@ -34,8 +34,8 @@ predictions <- predict(expeditions_model, newdata = new_data, type = "response")
 
 # Create a data frame to display the predictions
 prediction_df <- data.frame(
-  sex = new_data$sex, 
-  seasons = new_data$seasons, 
+  sex = new_data$sex,
+  seasons = new_data$seasons,
   height_range = new_data$height_range,
   age_range = new_data$age_range,
   solo = new_data$solo,
@@ -44,4 +44,3 @@ prediction_df <- data.frame(
 
 #### Saving the predictions ####
 write.csv(prediction_df, "data/predictions/predictions.csv", row.names = FALSE)
-

### 2024-04-18 paper/paper.qmd
Some grammatical errors fixed

@@
 Data from Himalayan expeditions in Nepal from 1905 through Spring 2019 is used and a Bayesian Logistic Regression model is leveraged to analyse the trends and factors influencing a successful summit.
-We find a strong relations between height of the peak, sex, age, season of expedition, if it was a solo ascent, and the success in summitting.
+I find a strong relation between height of the peak, sex, age, season of expedition, if it was a solo ascent, and the success in summitting.
 This research provides insights into factors affecting the successfullness of an expedition to a summit and helps for future expedition planning, risk management, and safety protocols."
@@
 Everest.
-Our paper uses the data from expeditions to the Himalayan mountain ranges in Nepal.
+My paper uses the data from expeditions to the Himalayan mountain ranges in Nepal.
 
-Our estimand is the relationship between different demographic, environmental and geographic factors (such as height of the peak, sex, age, season of expedition and if it was a solo ascent) and the successfulness of climbing a summit.
-Using our extensive analysis dataset, our goal is to identify trends and factors that influence a successful expedition and eventually conclude what factors help in a more successful attempt.
+My estimand is the relationship between different demographic, environmental and geographic factors (such as height of the peak, sex, age, season of expedition and if it was a solo ascent) and the successfulness of climbing a summit.
+Using the analysis dataset, my goal is to identify trends and factors that influence a successful expedition and eventually conclude what factors help in a more successful attempt.
 
-We use data from Alex Cookson’s datasets (link) which are sourced from The Himalayan Expedition records (link), to understand these factors and trends.
+I use data from Alex Cookson’s datasets (link) which are sourced from The Himalayan Expedition records (link), to understand these factors and trends.
 This is done by leveraging a Bayesian Logistic Regression model and then predicting the probability of a successful attempt over various demographic and environmental factors.
 
-Our analysis led to us … (add what was found)
+My analysis led to … (add what was found)
 
-This research highlights the interplay between climber attributes and environmental conditions, offering insights into Himalayan expeditions.
-These findings have practical implications for expedition planning, risk management, and safety protocols.
-Ultimately, our study aims to improve decision-making in high-altitude mountaineering, making it safer and more informed in one of the world's most challenging environments.
+The findings highlighted by this research has practical implications for expedition planning, risk management, and safety protocols.
+Ultimately, the study aims to improve decision-making in high-altitude mountaineering, making it safer and more informed in one of the world's most challenging environments.
 
@@
 @sec-data discusses how the dataset to be used for the analysis was obtained and pre-processed.
-I will explain the variables of interest in the dataset for the analysis.
+I will explain the variables of interest in the dataset used for the analysis.
 @sec-model describes the model being used for the analysis.
@@
 # Data {#sec-data}
-For this analysis, we have used combined three datasets into one, which is used for analysis.
+For this analysis, I have combined three datasets into one, which is used for analysis.
 The datasets were cleaned and analysed using the statistical programming software `R` [@citeR] along with the help `tidyverse` [@citeTidyverse], `knitr` [@citeKnitr], `ggplot2` [@citeGgplot], `here` [@citeHere], `dplyr` [@citeDplyr], `rstanarm` [@citeRstanarm], `arrow` [@citeArrow], `broom.mixed` [@citeBroomMixed], `modelsummary` [@citeModelSummary] and `kableExtra` [@citeKableExtra].
@@
 The three datasets I considered included information about all peaks, all expeditions on those peaks and all members on those expeditions.
-The data from these three datasets are combined to form our analysis dataset.
+The data from these three datasets are combined to form the main analysis dataset.
 
@@
 
-Among the overall range of variables available, we chose the following to be included in our analysis dataset.
+Among the overall range of variables available, I chose the following to be included in the analysis dataset.
 
-- `Height of the peak` in metres for the peak the person in the current entry is on an expedition for
+- `Height` is the height range in which the peak's heigh in metres falls.
+This is for the peak the person in the current entry is on an expedition for.
+The categoies for this are 5400 - 5749, 5750 - 6099, 6100 - 6449, 6450 - 6799, 6800 - 7149, 7150 - 7499, 7500 - 7849, 7850 - 8199, 8200 - 8549 and 8550 - 8900.
 - `Seasons` is the season the expedition is in. This takes on either of the 4 values: Autumn, Spring, Winter, Summer.
 - `Sex` is the sex reported by the expedition member and it is either male or female.
-- `Age` is the age of the expedition member at the time of the expedition.
+- `Age` is the age group in which the expedition member fell in at the time of the expedition.
 Depending on the best available data, this could be as of the summit date, the date of death, or the date of arrival at basecamp.
+The different categories for this are Under 18,19-30, 31-40, 41-50, 51-60, 61-70, 71-80 and 81-90.
 - `Success`represents whether the person was successful in summitting the goal peak.
@@
 
-Out of these, we will be using `height of the peak`, `seasons`, `sex`, `age` and `solo` in our model as the independent variables and `success` as the dependent variable.
+Out of these, I will be using `height of the peak`, `seasons`, `sex`, `age` and `solo` in the model as the independent variables and `success` as the dependent variable.
 
@@
 ## Cleaning
-For the analysis data, the cleaning steps we took were:
+For the analysis data, the cleaning steps I took were:
 

### 2024-04-18 paper/paper.qmd
Plots for data section added

@@
 
-(Plot the variables)
+```{r}
+#| echo: false
+#| warning: false
+#| message: false
+#| fig-cap: Counts for the variables of interest
+#| fig-subcap: ["Sex", "Solo Status", "Age", "Seasons", "Height of the Peak"]
+#| label: fig-counts
+#| layout-ncol: 2
+
+plot_bar <- function(data, x, x_label) {
+  if (x == "age_range") {
+    custom_order <- c("Under 18", "19-30", "31-40", "41-50", "51-60", "61-70", "71-80", "81-90")
+    data[[x]] <- factor(data[[x]], levels = custom_order)
+  }
+
+  ggplot(data, aes_string(x = x)) +
+    geom_bar(fill = "lightblue") +
+    labs(x = x_label, y = "Count") +
+    theme_minimal() +
+    theme(axis.text.x = element_text(angle = 45, hjust = 1))
+}
+
+plot_bar(analysis_data, "sex", "Sex")
+plot_bar(analysis_data, "solo", "Solo Status")
+plot_bar(analysis_data, "age_range", "Age")
+plot_bar(analysis_data, "seasons", "Season")
+plot_bar(analysis_data, "height_range", "Height of the Peak")
+
+```
+
+@fig-counts shows
+
+```{r}
+#| echo: false
+#| warning: false
+#| message: false
+#| fig-cap: Proportions for the variables of interest and the outcome of their expedition
+#| fig-subcap: ["Sex", "Solo Status", "Age", "Seasons", "Height of the Peak"]
+#| label: fig-proportions
+#| layout-ncol: 2
+
+library(tidyr)
+
+# Define a function to create the plot for each variable
+create_plot <- function(data, x_var, x_label) {
+  custom_order <- c("Under 18", "19-30", "31-40", "41-50", "51-60", "61-70", "71-80", "81-90")
+
+  # Create a dataframe with all possible combinations of x_var, success, and died
+  all_combinations <- expand.grid(
+    x_var = unique(data[[x_var]]),
+    success = c(TRUE, FALSE),
+    died = c(TRUE, FALSE)
+  )
+
+  # Join the original data with all_combinations to fill in missing combinations
+  plot_data <- all_combinations %>%
+    left_join(data, by = c(x_var = x_var, success = "success", died = "died")) %>%
+    group_by(x_var, success, died) %>%
+    summarise(count = n(), .groups = "drop") %>%
+    group_by(x_var) %>%
+    mutate(proportion = count / sum(count)) %>%
+    filter(success | died)
+
+  # If x_var is "age_range", convert it to a factor with custom order
+  if (x_var == "age_range") {
+    plot_data$x_var <- factor(plot_data$x_var, levels = custom_order)
+  }
+
+  p <- ggplot(plot_data, aes(x = x_var, y = proportion, fill = factor(success))) +
+    geom_bar(stat = "identity", position = "dodge") +
+    scale_fill_manual(values = c("TRUE" = "green", "FALSE" = "red"), 
+                      labels = c("Success", "Died"),
+                      name = "Outcome") +
+    labs(x = x_label, y = "Proportion") +
+    theme_minimal() +
+    theme(axis.text.x = element_text(angle = 45, hjust = 1))
+
+  return(p)
+}
+
+# Create plots for each variable
+create_plot(analysis_data, "sex", "Sex")
+create_plot(analysis_data, "solo", "Solo Status")
+create_plot(analysis_data, "age_range", "Age")
+create_plot(analysis_data, "seasons", "Season")
+create_plot(analysis_data, "height_range", "Height of the Peak")
+```
+
+@fig-proportions shows
 
 Out of these, I will be using `height of the peak`, `seasons`, `sex`, `age` and `solo` in the model as the independent variables and `success` as the dependent variable.
+
+\newpage
 
@@
 #| message: false
-#| label: fig-results
-#| fig-cap: Results1
+#| label: fig-sex
+#| fig-cap: Predicted Success Probability by Sex
 
@@
   geom_bar(stat = "identity", position = "dodge") +
-  labs(title = "Predicted Success Probability by Gender",
+  labs(title = "Predicted Success Probability by Sex",
        x = "Sex",
+       y = "Success Probability") +
+  theme_minimal()
+```
+```{r}
+#| echo: false
+#| warning: false
+#| message: false
+#| label: fig-age
+#| fig-cap: Predicted Success Probability by Age
+
+# Define custom order for age_range levels
+custom_order <- c("Under 18", "19-30", "31-40", "41-50", "51-60", "61-70", "71-80", "81-90")
+
+# Convert age_range to factor with custom order
+predictions$age_range <- factor(predictions$age_range, levels = custom_order)
+
+ggplot(predictions, aes(x = age_range, y = success_probability, fill = age_range)) +
+  geom_bar(stat = "identity", position = "dodge") +
+  labs(title = "Predicted Success Probability by Age",
+       x = "Age",
        y = "Success Probability") +
@@
 ```
+```{r}
+#| echo: false
+#| warning: false
+#| message: false
+#| label: fig-subplots
+#| fig-cap: Predicted Success Probability by Age and Sex
 
+# Define custom order for age_range levels
+custom_order <- c("Under 18", "19-30", "31-40", "41-50", "51-60", "61-70", "71-80", "81-90")
+
+# Convert age_range to factor with custom order
+predictions$age_range <- factor(predictions$age_range, levels = custom_order)
+
+ggplot(predictions, aes(x = age_range, y = success_probability, fill = sex)) +
+  geom_bar(stat = "identity", position = "dodge") +
+  labs(title = "Predicted Success Probability by Age and Sex",
+       x = "Age",
+       y = "Success Probability") +
+  facet_grid(sex ~ .) +  # Create subplots for each sex
+  theme_minimal()
+
+```
+
+```{r}
+#| echo: false
+#| warning: false
+#| message: false
+#| label: fig-subplots-height
+#| fig-cap: Predicted Success Probability by Height Range and Seasons
+
+ggplot(predictions, aes(x = height_range, y = success_probability, fill = seasons)) +
+  geom_bar(stat = "identity", position = "dodge") +
+  labs(title = "Predicted Success Probability by Height Range and Seasons",
+       x = "Height Range",
+       y = "Success Probability") +
+  facet_grid(seasons ~ .) +  # Create subplots for each season
+  theme_minimal()
+
+```
+
+```{r}
+#| echo: false
+#| warning: false
+#| message: false
+#| label: fig-subplots-sex-solo
+#| fig-cap: Predicted Success Probability by Sex and Solo Status
+
+ggplot(predictions, aes(x = sex, y = success_probability, fill = solo)) +
+  geom_bar(stat = "identity", position = "dodge") +
+  labs(title = "Predicted Success Probability by Sex and Solo Status",
+       x = "Sex",
+       y = "Success Probability") +
+  facet_grid(solo ~ .) +  # Create subplots for each solo status
+  theme_minimal()
+```
 # Discussion {#sec-discussion}

### 2024-04-18 paper/paper.qmd
Data section completed

@@
 The data from these three datasets are combined to form the main analysis dataset.
-
 A person becomes an entry in my analysis dataset if, between 1905 and Spring 2019, they attempted to climb any one of the many Himalayan peaks in Nepal.
@@
 
-- `Height` is the height range in which the peak's heigh in metres falls.
+- `Height` is the height range in which the peak's height in metres falls.
 This is for the peak the person in the current entry is on an expedition for.
-The categoies for this are 5400 - 5749, 5750 - 6099, 6100 - 6449, 6450 - 6799, 6800 - 7149, 7150 - 7499, 7500 - 7849, 7850 - 8199, 8200 - 8549 and 8550 - 8900.
-- `Seasons` is the season the expedition is in. This takes on either of the 4 values: Autumn, Spring, Winter, Summer.
+The categories for this are 5400 - 5749, 5750 - 6099, 6100 - 6449, 6450 - 6799, 6800 - 7149, 7150 - 7499, 7500 - 7849, 7850 - 8199, 8200 - 8549 and 8550 - 8900.
+- `Seasons` is the season the expedition is embarked in. This takes on either of the four values: Autumn, Spring, Winter, Summer.
 - `Sex` is the sex reported by the expedition member and it is either male or female.
@@
 The different categories for this are Under 18,19-30, 31-40, 41-50, 51-60, 61-70, 71-80 and 81-90.
-- `Success`represents whether the person was successful in summitting the goal peak.
+- `Success`represents whether the person's expedition resulted in a successful summit.
 - `Solo` represents whether the person attempted a solo ascent.
@@
 
-@fig-counts shows
+@fig-counts shows the counts for different variables we are considering to model.
+We see that there are significantly higher men than women.
+This shows the existence of a gender imbalance skewed towards men in mountaineering in the Himalayas.
+We also see other trends like most of the expedition members choose to not do solo ascents, most of the members fall in the middle-aged range and the peak expedition seasons are the more pleasant autumn and spring compared to the extreme seasons like winter and summer.
+These difference in counts need to be kept in mind when analysing success proportions.
 
@@
 
+  plot_data$success <- factor(plot_data$success, levels = c(TRUE, FALSE))
+
   p <- ggplot(plot_data, aes(x = x_var, y = proportion, fill = factor(success))) +
@@
 
-@fig-proportions shows
-
-Out of these, I will be using `height of the peak`, `seasons`, `sex`, `age` and `solo` in the model as the independent variables and `success` as the dependent variable.
+@fig-proportions shows the proportion of expedition members who has a successful outcome or died during their expedition.
+For sex, we see more men had successful expeditions compared to women.
+This is interesting keeping in the mind the vast difference in the counts of these two sexes going on expeditions.
+It might spark an interesting discussion into the gender imbalance in mountaineering.
+Additionally, we see most of the people who go on expeditions solo have more deaths which might explain the difference in the counts of people attempting solo ascent versus not attempting a solo ascent which was observed earlier.
+Some other trend we observe are that success rates decline with age and winter season contributes to the most deaths.
+The success and death proportions based on the height are very varied and do not follow a specific trend.
+This goes to show that just the height of the peak alone doesn't define the probability of having a successful summit but there are other factors which go into it.
 
@@
 For my analysis, a logistic regression model will be used to model if the person will be successful on their attempt to summit the particular Himalayan peak.
-The model will be based on five independent demographic variables: `height of the peak`, `sex`, `age`, `seasons` and `solo`.
+The model will be based on five independent demographic variables: `height of the peak`, `sex`, `age`, `seasons` and `solo` and the dependent variable will be `success`.
 

### 2024-04-18 paper/paper.qmd
Appendix cleaning information updated

@@
 
-1. Initial merging: I first merge the raw data from the `expeditions` and `members` datasets based on a common identifier, `expedition_id`, to consolidate information about expedition participants.
+1. Initial merging: The raw data from the expeditions and members datasets is merged based on a common identifier, expedition_id, consolidating information about expedition participants.
 
@@
 5.
-Final dataset creation: The resulting dataset is further refined to include only relevant columns (`peak_id`, `height_metres`, `season.x`, `sex`, `age`, `success`, `solo`, `died`).
-Some column names are adjusted for clarity, such as `season.x` being renamed to `seasons`, and `height_metres` being renamed to `height`.
+New ranges: Height range categories are created based on predefined ranges in meters: 5400-5749, 5750-6099, 6100-6449, 6450-6799, 6800-7149, 7150-7499, 7500-7849, 7850-8199, 8200-8549, and 8550-8900.
+Age range categories are defined as follows: Under 18, 19-30, 31-40, 41-50, 51-60, 61-70, 71-80, 81-90, and 91 or older.
+
+5.
+Final dataset creation: `season.x` is renamed to `seasons`.
+and then the resulting dataset is further refined to include only relevant columns (`height_range`, `seasons`, `sex`, `age_range`, `success`, `solo`, `died`).
 

### 2024-04-18 scripts/02-data_cleaning.R

@@ -33,21 +33,21 @@ merged_data2 <- merge(merged_data1, peaks, by = "peak_id")
 
 main_data <- merged_data2 %>%
   filter(!is.na(sex), !is.na(age)) %>% # Filter out rows with missing values for sex or age
-  rename(height = height_metres, seasons = season.x) %>%
+  rename(seasons = season.x) %>%
   mutate(
     # Create height range categories
     height_range = case_when(
-      height >= 5400 & height < 5750 ~ "5400 - 5749",
-      height >= 5750 & height < 6100 ~ "5750 - 6099",
-      height >= 6100 & height < 6450 ~ "6100 - 6449",
-      height >= 6450 & height < 6800 ~ "6450 - 6799",
-      height >= 6800 & height < 7150 ~ "6800 - 7149",
-      height >= 7150 & height < 7500 ~ "7150 - 7499",
-      height >= 7500 & height < 7850 ~ "7500 - 7849",
-      height >= 7850 & height < 8200 ~ "7850 - 8199",
-      height >= 8200 & height < 8550 ~ "8200 - 8549",
-      height >= 8550 & height <= 8900 ~ "8550 - 8900",
-      TRUE ~ as.character(height)
+      height_metres >= 5400 & height < 5750 ~ "5400 - 5749",
+      height >= 5750 & height_metres < 6100 ~ "5750 - 6099",
+      height_metres >= 6100 & height_metres < 6450 ~ "6100 - 6449",
+      height_metres >= 6450 & height_metres < 6800 ~ "6450 - 6799",
+      height_metres >= 6800 & height_metres < 7150 ~ "6800 - 7149",
+      height_metres >= 7150 & height_metres < 7500 ~ "7150 - 7499",
+      height_metres >= 7500 & height_metres < 7850 ~ "7500 - 7849",
+      height_metres >= 7850 & height_metres < 8200 ~ "7850 - 8199",
+      height_metres >= 8200 & height_metres < 8550 ~ "8200 - 8549",
+      height_metres >= 8550 & height_metres <= 8900 ~ "8550 - 8900",
+      TRUE ~ as.character(height_metres)
     ),
     # Create age range categories
     age_range = case_when(

### 2024-04-18 paper/paper.qmd
title and abstract modified, results section completed

@@
 ---
-title: "Himalayan Expeditions Success Analysis"
+title: "Understanding Factors Contributing to a Successful Himalayan Expedition"
 subtitle: "Analysis using data from Himalayan expeditions from 1905 through Spring 2019 leveraging Bayesian Logistic Regression"
@@
 date-format: long
-abstract: "I analyse the relationship between different demographic, environmental and geographic factors and the successfulness of an attempt to summit a peak.
+abstract: "This study analyses the relationship between different demographic, environmental and geographic factors and the successfulness of an attempt to summit a peak.
 Data from Himalayan expeditions in Nepal from 1905 through Spring 2019 is used and a Bayesian Logistic Regression model is leveraged to analyse the trends and factors influencing a successful summit.
 I find a strong relation between height of the peak, sex, age, season of expedition, if it was a solo ascent, and the success in summitting.
-This research provides insights into factors affecting the successfullness of an expedition to a summit and helps for future expedition planning, risk management, and safety protocols."
+Young age, being a man and embarking on the expedition in Spring or Autumn are some factors which increase your chances of having a successful summit.
+The insights from this study aim to help future expedition planning, risk management, and safety protocols."
 format: pdf
@@
 
-For my analysis, a logistic regression model will be used to model if the person will be successful on their attempt to summit the particular Himalayan peak.
-The model will be based on five independent demographic variables: `height of the peak`, `sex`, `age`, `seasons` and `solo` and the dependent variable will be `success`.
+My model will be based on five independent demographic variables: `height of the peak`, `sex`, `age`, `seasons` and `solo` and the dependent variable will be `success`.
 
@@
 
-- $\hat{p}$ represents the probability that someone will successfully complete the peak they are on the expedition for.
+- $\hat{p}$ represents the probability that someone will successfully summit the peak they are on the expedition for.
 - $\beta_0$ represents the intercept term of this logistical regression.
-It is the probability that someone will successfully complete the peak they are on the expedition for if the predictors' values are zero
+It is the probability that someone will successfully summit the peak they are on the expedition for if the predictors' values are zero
 - $\beta_1$ is the coefficient corresponding to height of the peak
@@
 
+# Results {#sec-results}
+
 ```{r}
@@
 ```
+
 @fig-model-coefficients illustrates the coefficients and their associated 95% credible intervals for the predictor variables in the Bayesian model.
@@
 These insights can help in understanding the direction and magnitude of the relationships between predictor variables and the outcome in the Bayesian model.
+I removed the coefficient for age_range81-90 which according to @tbl-model-coefficients was very small leading to all other coefficients being uninterpretable in the plot.
 
-# Results {#sec-results}
+We see that the males have a slightly higher success probability compared to females.
+We also notice how expeditions in Summer and Winter have lesser success probability than Autumn whereas Spring has higher success probability than it.
+Additionally, we notice the success probabilities getting lower with increase in age.
 
@@
 #| message: false
-#| label: fig-sex
-#| fig-cap: Predicted Success Probability by Sex
-
-ggplot(predictions, aes(x = sex, y = success_probability, fill = sex)) +
-  geom_bar(stat = "identity", position = "dodge") +
-  labs(title = "Predicted Success Probability by Sex",
-       x = "Sex",
-       y = "Success Probability") +
-  theme_minimal()
-```
-```{r}
-#| echo: false
-#| warning: false
-#| message: false
-#| label: fig-age
-#| fig-cap: Predicted Success Probability by Age
-
-# Define custom order for age_range levels
-custom_order <- c("Under 18", "19-30", "31-40", "41-50", "51-60", "61-70", "71-80", "81-90")
-
-# Convert age_range to factor with custom order
-predictions$age_range <- factor(predictions$age_range, levels = custom_order)
-
-ggplot(predictions, aes(x = age_range, y = success_probability, fill = age_range)) +
-  geom_bar(stat = "identity", position = "dodge") +
-  labs(title = "Predicted Success Probability by Age",
-       x = "Age",
-       y = "Success Probability") +
-  theme_minimal()
-
-```
-```{r}
-#| echo: false
-#| warning: false
-#| message: false
-#| label: fig-subplots
+#| label: fig-age-sex
 #| fig-cap: Predicted Success Probability by Age and Sex
@@
   geom_bar(stat = "identity", position = "dodge") +
-  labs(title = "Predicted Success Probability by Age and Sex",
-       x = "Age",
+  labs(x = "Age",
        y = "Success Probability") +
-  facet_grid(sex ~ .) +  # Create subplots for each sex
+  facet_grid(sex ~ .) +
   theme_minimal()
@@
 ```
+@fig-age-sex shows the predicted success probability according the the age and sex of the expedition member.
+We notice that the success probabilities over all ages tend to be slightly higher in males compared to females, just as we saw in @fig-model-coefficients.
+But, when combined with age we see there differences are not constant.
+We see that the difference in the success probabilities for people under 50 are very close to each other but for people aged 51 or above the difference is much larger.
+The success probabilities are pretty high, crossing 0.8 for younger individuals but starts declining with increase in age.
+The steep decline in category 81-90 might stem from the significant less number of expedition members in that age group leading to a less varied result.
 
@@
 #| message: false
-#| label: fig-subplots-height
-#| fig-cap: Predicted Success Probability by Height Range and Seasons
+#| label: fig-height-seasons
+#| fig-cap: Predicted Success Probability by Height of the Peak and Seasons
 
@@
   geom_bar(stat = "identity", position = "dodge") +
-  labs(title = "Predicted Success Probability by Height Range and Seasons",
-       x = "Height Range",
+  labs(x = "Height of the Peak",
        y = "Success Probability") +
-  facet_grid(seasons ~ .) +  # Create subplots for each season
-  theme_minimal()
+  facet_grid(seasons ~ .) + 
+  theme_minimal() +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))
 
 ```
+@fig-height-seasons shows the predicted success probability according the the height of the peak being ascended and the season of expedition.
+We notice that the success probabilities are pretty high for shorter mountain ranges and become lesser with increase in height.
+This might have to do with the fact that less higher peaks could be comparatively easier to summit.
+This, however, cannot be established as the trend as we see a non-pattern being followed for heights 6800 metres and above.
+We see that winter season expeditions have lower success probabilities, probably owning to the extreme cold conditions at higher altitudes.
+We also notice interesting trend as an interaction of these two factors.
+One such trend would be the success probabilities during different seasons for peaks between 5400 metres and 5749 metres are all comparatively closer to each other than compared to peaks between 7500 metres and 7849 metres in height.
+All seasons' success probabilities for peaks between 5400 metres and 5749 metres fall between 0.85 to 0.9 whereas for for peaks between 7500 metres and 7848 metres, the success probabilities go from around 0.35 in Winter to around 0.6 in Autumn or Spring.
 
-```{r}
-#| echo: false
-#| warning: false
-#| message: false
-#| label: fig-subplots-sex-solo
-#| fig-cap: Predicted Success Probability by Sex and Solo Status
-
-ggplot(predictions, aes(x = sex, y = success_probability, fill = solo)) +
-  geom_bar(stat = "identity", position = "dodge") +
-  labs(title = "Predicted Success Probability by Sex and Solo Status",
-       x = "Sex",
-       y = "Success Probability") +
-  facet_grid(solo ~ .) +  # Create subplots for each solo status
-  theme_minimal()
-```
 # Discussion {#sec-discussion}
@@
 
+## Model summary
+
+## Hello
+```{r}
+#| label: fig-post_dist
+#| fig-cap: Posterior distribution for logistic regression model 
+#| echo: false
+#| warning: false
+#| message: false
+
+library(bayesplot)
+
+pp_check(expeditions_model) +
+  theme(legend.position = "bottom")
+```
+
 # References

### 2024-04-18 paper/paper.qmd
Introduction completed

@@
 I find a strong relation between height of the peak, sex, age, season of expedition, if it was a solo ascent, and the success in summitting.
-Young age, being a man and embarking on the expedition in Spring or Autumn are some factors which increase your chances of having a successful summit.
+Young age, being a man and embarking on the expedition in Spring or Autumn are some factors which increase one's chances of having a successful summit.
 The insights from this study aim to help future expedition planning, risk management, and safety protocols."
@@
 
-My analysis led to … (add what was found)
+My analysis shows how mountaineering is a highly male dominated activity.
+It highlights that young age, being a man and embarking on the expedition in Spring or Autumn are some factors which increase one's chances of having a successful summit.
 
-The findings highlighted by this research has practical implications for expedition planning, risk management, and safety protocols.
+The findings highlighted by this research have practical implications for expedition planning, risk management, and safety protocols.
 Ultimately, the study aims to improve decision-making in high-altitude mountaineering, making it safer and more informed in one of the world's most challenging environments.

### 2024-04-18 other/datasheet/datasheet.qmd
Datasheet for analysis data added

---
title: "Datasheet for Analysis Data: analysis_data.parquet"
subtitle: "Dataset can be found  at data/analysis_data"
author: Kaavya Kalani
thanks: "Code and data are available at: [https://github.com/kaavyakalani26/himalayan-expeditions-analysis](https://github.com/kaavyakalani26/himalayan-expeditions-analysis)"
date: today
date-format: long
format: pdf
number-sections: true
bibliography: references.bib
---

Extract of the questions from @gebru2021datasheets.

**Motivation**

1. *For what purpose was the dataset created? Was there a specific task in mind? Was there a specific gap that needed to be filled? Please provide a description.*
    - The dataset was created for the analysis of Himalayan expeditions. The analysis was of factors affecting a successful summit. There are multiple datasets with different forms of information about expeditions available but no singular dataset that contained all the variables the analysis was looking for.

2. *Who created the dataset (for example, which team, research group) and on behalf of which entity (for example, company, institution, organization)?*
    - Kaavya Kalani, while studying at the University of Toronto.

3. *Who funded the creation of the dataset? If there is an associated grant, please provide the name of the grantor and the grant name and number.*
    - No funding was received for this project.

4. *Any other comments?*
    - No.

\newpage

**Composition**

1. *What do the instances that comprise the dataset represent (for example, documents, photos, people, countries)? Are there multiple types of instances (for example, movies, users, and ratings; people and interactions between them; nodes and edges)? Please provide a description.*
	- Each row of the dataset represents information about an individual on an expedition on a peak in the Nepal Himalaya.

2. *How many instances are there in total (of each type, if appropriate)?*
	- 73014

3. *Does the dataset contain all possible instances or is it a sample (not necessarily random) of instances from a larger set? If the dataset is a sample, then what is the larger set? Is the sample representative of the larger set (for example, geographic coverage)? If so, please describe how this representativeness was validated/verified. If it is not representative of the larger set, please describe why not (for example, to cover a more diverse range of instances, because instances were withheld or unavailable).*
	- The dataset contains contains all instances from 1905 through Spring 2019.

4. *What data does each instance consist of? "Raw" data (for example, unprocessed text or images) or features? In either case, please provide a description.*
	- Each instance consists of demographic information about the person in terms of their age and sex, it contains information about the height of the peak the individual is on an expedition for, the season of the expedition, the ascent was attempted solo, if the ascent was successful and if the person died during the expedition.

5. *Is there a label or target associated with each instance? If so, please provide a description.*
	- No.

6. *Is any information missing from individual instances? If so, please provide a description, explaining why this information is missing (for example, because it was unavailable). This does not include intentionally removed information, but might include, for example, redacted text.*
	- No.

7. *Are relationships between individual instances made explicit (for example, users' movie ratings, social network links)? If so, please describe how these relationships are made explicit.*
	- No, as there are no relationships between different instances.

8. *Are there recommended data splits (for example, training, development/validation, testing)? If so, please provide a description of these splits, explaining the rationale behind them.*
	- No.

9. *Are there any errors, sources of noise, or redundancies in the dataset? If so, please provide a description.*
	- No.

10. *Is the dataset self-contained, or does it link to or otherwise rely on external resources (for example, websites, tweets, other datasets)? If it links to or relies on external resources, a) are there guarantees that they will exist, and remain constant, over time; b) are there official archival versions of the complete dataset (that is, including the external resources as they existed at the time the dataset was created); c) are there any restrictions (for example, licenses, fees) associated with any of the external resources that might apply to a dataset consumer? Please provide descriptions of all external resources and any restrictions associated with them, as well as links or other access points, as appropriate.*
	- Self-contained.

11. *Does the dataset contain data that might be considered confidential (for example, data that is protected by legal privilege or by doctor-patient confidentiality, data that includes the content of individuals' non-public communications)? If so, please provide a description.*
	- No, all data were gathered from public sources.

12. *Does the dataset contain data that, if viewed directly, might be offensive, insulting, threatening, or might otherwise cause anxiety? If so, please describe why.*
	- No.

13. *Does the dataset identify any sub-populations (for example, by age, gender)? If so, please describe how these subpopulations are identified and provide a description of their respective distributions within the dataset.*
	- Yes, age and sex.

14. *Is it possible to identify individuals (that is, one or more natural persons), either directly or indirectly (that is, in combination with other data) from the dataset? If so, please describe how.*
	- No, atleast no other data known of.

\newpage

15. *Does the dataset contain data that might be considered sensitive in any way (for example, data that reveals race or ethnic origins, sexual orientations, religious beliefs, political opinions or union memberships, or locations; financial or health data; biometric or genetic data; forms of government identification, such as social security numbers; criminal history)? If so, please provide a description.*
	- No.

16. *Any other comments?*
	- No.

**Collection process**

1. *How was the data associated with each instance acquired? Was the data directly observable (for example, raw text, movie ratings), reported by subjects (for example, survey responses), or indirectly inferred/derived from other data (for example, part-of-speech tags, model-based guesses for age or language)? If the data was reported by subjects or indirectly inferred/derived from other data, was the data validated/verified? If so, please describe how.*
	- The data was obtained by cleaning the dataset's provided by Alex Cookson: (link). Their datasets were derived from The Himalayan Database (link). The Himalayan Database is a compilation of records for all expeditions that have climbed in the Nepal Himalaya. The database is based on the expedition archives of Elizabeth Hawley, a longtime journalist based in Kathmandu, and it is supplemented by information gathered from books, alpine journals and correspondence with Himalayan climbers.

2. *What mechanisms or procedures were used to collect the data (for example, hardware apparatuses or sensors, manual human curation, software programs, software APIs)? How were these mechanisms or procedures validated?*
	- The csv files provided by Alex Cookson (link) were downloaded and processed. The cleaning and processing involved natural joining the three datasets and cleaning them to satisfy the analysis needs.

3. *If the dataset is a sample from a larger set, what was the sampling strategy (for example, deterministic, probabilistic with specific sampling probabilities)?*
	- Dataset is not a sample.

4. *Who was involved in the data collection process (for example, students, crowdworkers, contractors) and how were they compensated (for example, how much were crowdworkers paid)?*
	- Kaavya Kalani, and this was done for no compensation.

5. *Over what timeframe was the data collected? Does this timeframe match the creation timeframe of the data associated with the instances (for example, recent crawl of old news articles)? If not, please describe the timeframe in which the data associated with the instances was created.*
	- The dataset contain instances from 1905 through Spring 2019.

6. *Were any ethical review processes conducted (for example, by an institutional review board)? If so, please provide a description of these review processes, including the outcomes, as well as a link or other access point to any supporting documentation.*
	- No.

7. *Did you collect the data from the individuals in question directly, or obtain it via third parties or other sources (for example, websites)?*
	- Third party sources.

8. *Were the individuals in question notified about the data collection? If so, please describe (or show with screenshots or other information) how notice was provided, and provide a link or other access point to, or otherwise reproduce, the exact language of the notification itself.*
	- No.

9. *Did the individuals in question consent to the collection and use of their data? If so, please describe (or show with screenshots or other information) how consent was requested and provided, and provide a link or other access point to, or otherwise reproduce, the exact language to which the individuals consented.*
	- Technically yes. The repository from which the data was extracted states, "Feel free to use this data, but none of it belongs to me, so please do so in accordance with the dataset's license. You don't need to cite me if you use something I scraped or cleaned. (But I would love to hear about what you're doing with it!)"

10. *If consent was obtained, were the consenting individuals provided with a mechanism to revoke their consent in the future or for certain uses? If so, please provide a description, as well as a link or other access point to the mechanism (if appropriate).*
	- No, as it was free to use data.

11. *Has an analysis of the potential impact of the dataset and its use on data subjects (for example, a data protection impact analysis) been conducted? If so, please provide a description of this analysis, including the outcomes, as well as a link or other access point to any supporting documentation.*
	- No.

\newpage

12. *Any other comments?*
	- No.

**Preprocessing/cleaning/labeling**

1. *Was any preprocessing/cleaning/labeling of the data done (for example, discretization or bucketing, tokenization, part-of-speech tagging, SIFT feature extraction, removal of instances, processing of missing values)? If so, please provide a description. If not, you may skip the remaining questions in this section.*
	- Yes.

2. *Was the "raw" data saved in addition to the preprocessed/cleaned/labeled data (for example, to support unanticipated future uses)? If so, please provide a link or other access point to the "raw" data.*
	- Yes. `data/raw_data` contains the raw data.

3. *Is the software that was used to preprocess/clean/label the data available? If so, please provide a link or other access point.*
	- R was used.

4. *Any other comments?*
	- No.

**Uses**

1. *Has the dataset been used for any tasks already? If so, please provide a description.*
	- The dataset has been used by Kaavya Kalani for the analysing the factors contributing to a successful Himalayan peak summit. No other use of the dataset is known.

2. *Is there a repository that links to any or all papers or systems that use the dataset? If so, please provide a link or other access point.*
	- No.

3. *What (other) tasks could the dataset be used for?*
	- Linking with other demographic and environmental factors would be interesting.

4. *Is there anything about the composition of the dataset or the way it was collected and preprocessed/cleaned/labeled that might impact future uses? For example, is there anything that a dataset consumer might need to know to avoid uses that could result in unfair treatment of individuals or groups (for example, stereotyping, quality of service issues) or other risks or harms (for example, legal risks, financial harms)? If so, please provide a description. Is there anything a dataset consumer could do to mitigate these risks or harms?*
	- No.

5. *Are there tasks for which the dataset should not be used? If so, please provide a description.*
	- No.

6. *Any other comments?*
	- No.

**Distribution**

1. *Will the dataset be distributed to third parties outside of the entity (for example, company, institution, organization) on behalf of which the dataset was created? If so, please provide a description.*
	- The dataset is available through GitHub.

2. *How will the dataset be distributed (for example, tarball on website, API, GitHub)? Does the dataset have a digital object identifier (DOI)?*
	- Github.

3. *When will the dataset be distributed?*
	- The dataset is available now.

4. *Will the dataset be distributed under a copyright or other intellectual property (IP) license, and/or under applicable terms of use (ToU)? If so, please describe this license and/ or ToU, and provide a link or other access point to, or otherwise reproduce, any relevant licensing terms or ToU, as well as any fees associated with these restrictions.*
	- No. MIT license.

5. *Have any third parties imposed IP-based or other restrictions on the data associated with the instances? If so, please describe these restrictions, and provide a link or other access point to, or otherwise reproduce, any relevant licensing terms, as well as any fees associated with these restrictions.*
	- None that are known.

6. *Do any export controls or other regulatory restrictions apply to the dataset or to individual instances? If so, please describe these restrictions, and provide a link or other access point to, or otherwise reproduce, any supporting documentation.*
	- None that are known.

7. *Any other comments?*
	- No.

**Maintenance**

1. *Who will be supporting/hosting/maintaining the dataset?*
	- Kaavya Kalani

2. *How can the owner/curator/manager of the dataset be contacted (for example, email address)?*
	- kaavya.kalani@mail.utoronto.ca

3. *Is there an erratum? If so, please provide a link or other access point.*
	- No.

4. *Will the dataset be updated (for example, to correct labeling errors, add new instances, delete instances)? If so, please describe how often, by whom, and how updates will be communicated to dataset consumers (for example, mailing list, GitHub)?*
	- No.

5. *If the dataset relates to people, are there applicable limits on the retention of the data associated with the instances (for example, were the individuals in question told that their data would be retained for a fixed period of time and then deleted)? If so, please describe these limits and explain how they will be enforced.*
	- No.

6. *Will older versions of the dataset continue to be supported/hosted/maintained? If so, please describe how. If not, please describe how its obsolescence will be communicated to dataset consumers.*
	- No, as the dataset is just updated. A history of the updates is available through GitHub.

7. *If others want to extend/augment/build on/contribute to the dataset, is there a mechanism for them to do so? If so, please provide a description. Will these contributions be validated/verified? If so, please describe how. If not, why not? Is there a process for communicating/distributing these contributions to dataset consumers? If so, please provide a description.*
	- Pull request on GitHub.

8. *Any other comments?*
	- No.

\newpage

# References

### 2024-04-18 paper/paper.qmd
styler ran on qmd

@@
 plot_bar(analysis_data, "height_range", "Height of the Peak")
-
 ```
@@
     geom_bar(stat = "identity", position = "dodge") +
-    scale_fill_manual(values = c("TRUE" = "green", "FALSE" = "red"), 
+    scale_fill_manual(
+      values = c("TRUE" = "green", "FALSE" = "red"),
       labels = c("Success", "Died"),
-                      name = "Outcome") +
+      name = "Outcome"
+    ) +
     labs(x = x_label, y = "Proportion") +
@@
   geom_bar(stat = "identity", position = "dodge") +
-  labs(x = "Age",
-       y = "Success Probability") +
+  labs(
+    x = "Age",
+    y = "Success Probability"
+  ) +
   facet_grid(sex ~ .) +
   theme_minimal()
-
 ```
@@
   geom_bar(stat = "identity", position = "dodge") +
-  labs(x = "Height of the Peak",
-       y = "Success Probability") +
+  labs(
+    x = "Height of the Peak",
+    y = "Success Probability"
+  ) +
   facet_grid(seasons ~ .) +
@@
   theme(axis.text.x = element_text(angle = 45, hjust = 1))
-
 ```

### 2024-04-19 paper/paper.qmd
Comments added

@@
 
+#### Workspace setup ####
 library(broom.mixed)
@@
 
+#### Load necessary data and models ####
 expeditions_model <- readRDS(here("models/single_bay.rds"))
@@
 
+# Function to generate the plots
 plot_bar <- function(data, x, x_label) {
+  # Check if x is "age_range" to customize the order of bars
   if (x == "age_range") {
@@
 
+  # Create a bar plot 
   ggplot(data, aes_string(x = x)) +
@@
 
+# Create plots for each variable 
 plot_bar(analysis_data, "sex", "Sex")
@@
 
-library(tidyr)
-
-# Define a function to create the plot for each variable
+# Function to create the plot for each variable
 create_plot <- function(data, x_var, x_label) {
@@
 #| tbl-cap: Summary of the model
-#
-# # Create model summary
-# model_summary <- modelsummary(expeditions_model)
 
@@
 
+# Remove the coefficient for age_range81-90 as it is very small
 coefficients <- coefficients[coefficients$term != "age_range81-90", ]
 
+# Plot for the coefficients of the model
 ggplot(coefficients, aes(x = estimate, y = term)) +
@@
 
+# Plot for the predicted success probability by age and sex
 ggplot(predictions, aes(x = age_range, y = success_probability, fill = sex)) +
@@
 
+# Plot for the predicted success probability by height of the peak and season of expedition
 ggplot(predictions, aes(x = height_range, y = success_probability, fill = seasons)) +
@@
 #| tbl-cap: Analysis dataset
-#|
+
+# Glimpse of the anaysis dataset
 head(analysis_data) %>%
@@
 
-## Hello
+## Posterior distribution
 ```{r}

### 2024-04-19 other/datasheet/datasheet.qmd
Dataset related links and citations added

@@
 If so, please describe how.*
-	- The data was obtained by cleaning the dataset's provided by Alex Cookson: (link).
-Their datasets were derived from The Himalayan Database (link).
+	- The data was obtained by cleaning the dataset's provided on Github by @Cookson.
+Their datasets were derived from The Himalayan Database.
 The Himalayan Database is a compilation of records for all expeditions that have climbed in the Nepal Himalaya.
@@
 How were these mechanisms or procedures validated?*
-	- The csv files provided by Alex Cookson (link) were downloaded and processed. The cleaning and processing involved natural joining the three datasets and cleaning them to satisfy the analysis needs.
+	- The csv files provided by @Cookson were downloaded and processed. The cleaning and processing involved natural joining the three datasets and cleaning them to satisfy the analysis needs.
 

### 2024-04-19 paper/paper.qmd
Dataset related links and citations added

@@
 
-I use data from Alex Cookson’s datasets (link) which are sourced from The Himalayan Expedition records (link), to understand these factors and trends.
+I use data from @Cookson which are sourced from The Himalayan Expedition records (@Salisbury), to understand these factors and trends.
 This is done by leveraging a Bayesian Logistic Regression model and then predicting the probability of a successful attempt over various demographic and environmental factors.
@@
 ## Analysis Dataset
-The raw datasets were obtained from Alex Cookson's datasets (link). I chose the ones cleaned for Himalayan expeditions. Alex got his datasets from The Himalayan Database (link).
+The raw datasets were obtained from @Cookson. I chose the ones cleaned for Himalayan expeditions. Cookson got his datasets from The Himalayan Database (@Salisbury).
 
@@
 
-My dataset derived from Alex's contains the entries from 1905 through Spring 2019.
+My dataset derived from Cookson's contains the entries from 1905 through Spring 2019.
 

### 2024-04-19 paper/paper.qmd
Discussion added and language enhanced

@@
 date-format: long
-abstract: "This study analyses the relationship between different demographic, environmental and geographic factors and the successfulness of an attempt to summit a peak.
+abstract: "This study analyses the relationship between various demographic, environmental and geographic factors and the probability of success in attempting to summit a peak.
 Data from Himalayan expeditions in Nepal from 1905 through Spring 2019 is used and a Bayesian Logistic Regression model is leveraged to analyse the trends and factors influencing a successful summit.
-I find a strong relation between height of the peak, sex, age, season of expedition, if it was a solo ascent, and the success in summitting.
-Young age, being a man and embarking on the expedition in Spring or Autumn are some factors which increase one's chances of having a successful summit.
-The insights from this study aim to help future expedition planning, risk management, and safety protocols."
+Young age, being male and embarking on the expedition in spring or autumn are some factors that increase one's chances of a successful summit.
+The insights from this study aim to assist in future expedition planning, risk management, and safety protocols."
 format: pdf
@@
 Mountaineering, with its blend of adventure and challenge, has captivated explorers for generations.
-It represents a pinnacle of human endeavor, testing physical prowess, mental fortitude, and strategic planning against some of the most formidable natural landscapes on Earth.
-Amidst the allure of conquering these majestic peaks, lies a critical question: what factors contribute to the likelihood of a successful summit attempt?
-This paper delves into precisely this inquiry.
+It represents a pinnacle of human endeavor, testing physical strength, mental fortitude, and strategic planning against some of the most formidable natural landscapes on Earth.
+Amidst the allure of conquering majestic peaks, a critical question arises: what factors contribute to the likelihood of a successful summit attempt?
+This paper delves into precisely this.
 
-There are numerous mountain ranges around the world where people embark on expeditions.
-One of the most famous mountain ranges are the Himalayas, which also consist of the highest mountain peak in the world, Mt.
-Everest.
-My paper uses the data from expeditions to the Himalayan mountain ranges in Nepal.
+Numerous mountain ranges worldwide attract expedition enthusiasts, with the Himalayas, home to the highest peak, Mt.
+Everest, being one of the most renowned.
+This paper focuses on data from expeditions to the Himalayan mountain ranges in Nepal.
 
-My estimand is the relationship between different demographic, environmental and geographic factors (such as height of the peak, sex, age, season of expedition and if it was a solo ascent) and the successfulness of climbing a summit.
-Using the analysis dataset, my goal is to identify trends and factors that influence a successful expedition and eventually conclude what factors help in a more successful attempt.
+My estimand is the relationship between various demographic, environmental and geographic factors (such as the height of the peak, sex, age, season of the expedition and if it was a solo ascent) and the success of an expedition.
+By analysing the dataset, my goal is to identify trends and factors influencing successful expeditions and determine which factors contribute to a higher likelihood of success.
 
-I use data from @Cookson which are sourced from The Himalayan Expedition records (@Salisbury), to understand these factors and trends.
-This is done by leveraging a Bayesian Logistic Regression model and then predicting the probability of a successful attempt over various demographic and environmental factors.
+Data sourced from @Cookson, based on The Himalayan Expedition records [@Salisbury], is utilized  to understand these factors and trends.
+A Bayesian Logistic Regression model is employed to predict the probability of a successful attempt based on demographic, environmental and geographic factors.
 
-My analysis shows how mountaineering is a highly male dominated activity.
-It highlights that young age, being a man and embarking on the expedition in Spring or Autumn are some factors which increase one's chances of having a successful summit.
+My analysis reveals that mountaineering is a predominantly male-dominated activity, with factors such as young age, male gender, and expeditions in spring or autumn increasing the chances of having a successful summit.
 
-The findings highlighted by this research have practical implications for expedition planning, risk management, and safety protocols.
-Ultimately, the study aims to improve decision-making in high-altitude mountaineering, making it safer and more informed in one of the world's most challenging environments.
+The findings of this research have practical implications for expedition planning, risk management, and safety protocol aiming to enhance decision-making in high-altitude mountaineering, making it safer and more informed in one of the world's most challenging environments.
 
 The paper is further organized into four sections.
-@sec-data discusses how the dataset to be used for the analysis was obtained and pre-processed.
+@sec-data discusses how the dataset used for the analysis was obtained and pre-processed.
 I will explain the variables of interest in the dataset used for the analysis.
@@
 @sec-results then highlights and discusses the trends and associations found during the analysis.
-Lastly, @sec-discussion talks about some interesting trends found in @sec-results in depth, link it to the real world and also highlight the weaknesses and future of my analysis.
+Lastly, @sec-discussion talks about some interesting trends found in @sec-results in depth, linking it to the real world and also highlights the weaknesses and future of my analysis.
 
@@
 For this analysis, I have combined three datasets into one, which is used for analysis.
-The datasets were cleaned and analysed using the statistical programming software `R` [@citeR] along with the help `tidyverse` [@citeTidyverse], `knitr` [@citeKnitr], `ggplot2` [@citeGgplot], `here` [@citeHere], `dplyr` [@citeDplyr], `rstanarm` [@citeRstanarm], `arrow` [@citeArrow], `broom.mixed` [@citeBroomMixed], `modelsummary` [@citeModelSummary] and `kableExtra` [@citeKableExtra].
+The datasets were cleaned and analysed using the statistical programming software `R` [@citeR] along with the help of `tidyverse` [@citeTidyverse], `knitr` [@citeKnitr], `ggplot2` [@citeGgplot], `here` [@citeHere], `dplyr` [@citeDplyr], `rstanarm` [@citeRstanarm], `arrow` [@citeArrow], `broom.mixed` [@citeBroomMixed] and `kableExtra` [@citeKableExtra].
 
 ## Analysis Dataset
-The raw datasets were obtained from @Cookson. I chose the ones cleaned for Himalayan expeditions. Cookson got his datasets from The Himalayan Database (@Salisbury).
+The raw datasets were obtained from @Cookson, who sourced them from The Himalayan Database [@Salisbury].
 
@@
 
-My dataset derived from Cookson's contains the entries from 1905 through Spring 2019.
+My dataset, derived from Cookson's, contains the entries from 1905 through Spring 2019.
 
-The three datasets I considered included information about all peaks, all expeditions on those peaks and all members on those expeditions.
-The data from these three datasets are combined to form the main analysis dataset.
+The three datasets I considered included information about all peaks in Nepal, all expeditions on those peaks and all members on those expeditions.
+I then combines the data from these three datasets to form the main analysis dataset.
+
 A person becomes an entry in my analysis dataset if, between 1905 and Spring 2019, they attempted to climb any one of the many Himalayan peaks in Nepal.
@@
 The categories for this are 5400 - 5749, 5750 - 6099, 6100 - 6449, 6450 - 6799, 6800 - 7149, 7150 - 7499, 7500 - 7849, 7850 - 8199, 8200 - 8549 and 8550 - 8900.
-- `Seasons` is the season the expedition is embarked in. This takes on either of the four values: Autumn, Spring, Winter, Summer.
+- `Seasons` is the season the expedition is embarked on. This takes on either of the four values: autumn, spring, winter, summer.
 - `Sex` is the sex reported by the expedition member and it is either male or female.
-- `Age` is the age group in which the expedition member fell in at the time of the expedition.
+- `Age` is the age group in which the expedition member fell at the time of the expedition.
 Depending on the best available data, this could be as of the summit date, the date of death, or the date of arrival at basecamp.
-The different categories for this are Under 18,19-30, 31-40, 41-50, 51-60, 61-70, 71-80 and 81-90.
-- `Success`represents whether the person's expedition resulted in a successful summit.
-- `Solo` represents whether the person attempted a solo ascent.
-- `Died` represents whether the person died during the expedition.
+The different categories for this are Under 18, 19-30, 31-40, 41-50, 51-60, 61-70, 71-80 and 81-90.
+- `Success` indicates if the person's expedition resulted in a successful summit.
+- `Solo` indicates if the person attempted a solo ascent.
+- `Died` indicates if the person died during the expedition.
 
@@
 
-@fig-counts shows the counts for different variables we are considering to model.
+@fig-counts displays the counts for various variables I will be modelling.
 We see that there are significantly higher men than women.
 This shows the existence of a gender imbalance skewed towards men in mountaineering in the Himalayas.
-We also see other trends like most of the expedition members choose to not do solo ascents, most of the members fall in the middle-aged range and the peak expedition seasons are the more pleasant autumn and spring compared to the extreme seasons like winter and summer.
-These difference in counts need to be kept in mind when analysing success proportions.
+Additionally, most expedition members opt not to attempt solo ascents and fall within the middle-aged range.
+Lastly, we see the peak expedition seasons are the more pleasant autumn and spring over the extreme seasons like summer and winter.
+These count disparities should be considered when analysing success proportions.
 
@@
 
-@fig-proportions shows the proportion of expedition members who has a successful outcome or died during their expedition.
-For sex, we see more men had successful expeditions compared to women.
-This is interesting keeping in the mind the vast difference in the counts of these two sexes going on expeditions.
-It might spark an interesting discussion into the gender imbalance in mountaineering.
-Additionally, we see most of the people who go on expeditions solo have more deaths which might explain the difference in the counts of people attempting solo ascent versus not attempting a solo ascent which was observed earlier.
-Some other trend we observe are that success rates decline with age and winter season contributes to the most deaths.
-The success and death proportions based on the height are very varied and do not follow a specific trend.
-This goes to show that just the height of the peak alone doesn't define the probability of having a successful summit but there are other factors which go into it.
+@fig-proportions shows the proportion of expedition members who had a successful outcome or died during their expedition.
+For sex, we see that more men had successful expeditions compared to women.
+Moreover, individuals who embark on solo expeditions have a higher mortality rate, potentially explaining the difference in the number of people attempting solo ascent observed earlier.
+Some other trends that can be observed are that success rates decline with age, and winter contributes to the most deaths.
+The success and death proportions based on the height of the peak are very varied and do not follow a specific trend.
+This goes to show that just the height of the peak alone doesn't define the probability of having a successful summit but there are other factors that contribute to it.
 
@@
 # Model {#sec-model}
-I used a Bayesian Logistic Regression model to find the probability that someone will successfully summit the Himalayan peak they are on the expedition for.
-Logistic regression is a method used for binary classification to predict the probability of a categorical dependent variable.
+I used a Bayesian Logistic Regression model to determine the likelihood of an individual successfully reaching the summit of the Himalayan peak they are attempting to climb during the expedition.
+Logistic regression is a statistical technique used for binary classification to predict the probability of a categorical dependent variable.
 
-My model will be based on five independent demographic variables: `height of the peak`, `sex`, `age`, `seasons` and `solo` and the dependent variable will be `success`.
+My model will be based on five independent demographic variables: `height`, `sex`, `age`, `seasons` and `solo` and the dependent variable will be `success`.
 
@@
 
-```{r}
-#| echo: false
-#| warning: false
-#| message: false
-#| label: tbl-model-coefficients
-#| tbl-cap: Summary of the model
-
-# Extract coefficients
-coefficients <- broom::tidy(expeditions_model, conf.int = TRUE)
-coefficients |> kable(digits = 2)
-```
-@tbl-model-coefficients shows the coefficients for my Bayesian model along with the standard error and the 95% credible interval.
-The standard error (SE) is a measure of the precision with which a sample statistic estimates a population parameter.
-It quantifies the variability of sample statistics around the population parameter.
-A 95% credible interval means that there is a 95% probability that the true parameter lies within the interval, given the observed data and the model assumptions.
-
 # Results {#sec-results}
@@
 #| fig-cap: Coefficients of the model
+
+# Coefficients for the model
+coefficients <- broom::tidy(expeditions_model, conf.int = TRUE)
 
@@
 
-@fig-model-coefficients illustrates the coefficients and their associated 95% credible intervals for the predictor variables in the Bayesian model.
-Each point represents the estimated coefficient for a predictor variable, while the horizontal lines depict the credible interval around the estimate.
+@fig-model-coefficients illustrates the coefficients and their corresponding 95% credible intervals for the predictor variables in the Bayesian model.
+A 95% credible interval means that there is a 95% probability that the true parameter lies within the interval, given the observed data and the model assumptions.
+Each point represents the estimated coefficient for a predictor variable, with the horizontal lines indicating the credible interval around the estimate.
 Variables with coefficients to the right of zero indicate a positive association with the outcome variable, suggesting that an increase in the predictor variable corresponds to an increase in the outcome variable.
@@
 These insights can help in understanding the direction and magnitude of the relationships between predictor variables and the outcome in the Bayesian model.
-I removed the coefficient for age_range81-90 which according to @tbl-model-coefficients was very small leading to all other coefficients being uninterpretable in the plot.
+The coefficient for the age_range81-90 variable was removed due to its small magnitude, which was affecting the interpretability of other coefficients in the plot.
 
-We see that the males have a slightly higher success probability compared to females.
-We also notice how expeditions in Summer and Winter have lesser success probability than Autumn whereas Spring has higher success probability than it.
-Additionally, we notice the success probabilities getting lower with increase in age.
+The analysis shows that males have a slightly higher success probability compared to females.
+Furthermore, expeditions in summer and winter show lower success probability than autumn, while spring shows a higher success probability.
+Additionally, the success probabilities decrease with increase in age.
 
@@
 ```
-@fig-age-sex shows the predicted success probability according the the age and sex of the expedition member.
-We notice that the success probabilities over all ages tend to be slightly higher in males compared to females, just as we saw in @fig-model-coefficients.
-But, when combined with age we see there differences are not constant.
-We see that the difference in the success probabilities for people under 50 are very close to each other but for people aged 51 or above the difference is much larger.
-The success probabilities are pretty high, crossing 0.8 for younger individuals but starts declining with increase in age.
-The steep decline in category 81-90 might stem from the significant less number of expedition members in that age group leading to a less varied result.
+@fig-age-sex shows the predicted success probability based on the age and sex of expedition members.
+It is evident that success probabilities are slightly higher for males across all ages, consistent with the findings in @fig-model-coefficients.
+However, when age is factored in, these differences vary.
+For individuals under 50, the success probabilities for males and females are similar, but for those aged 51 and above, the gap widens significantly.
+Success probabilities are notably high, exceeding 0.8 for younger individuals but decrease with age.
+The sharp decline in the 81-90 age category may be attributed to the smaller number of expedition members in that age group, resulting in less reliable outcomes.
 
@@
 ```
-@fig-height-seasons shows the predicted success probability according the the height of the peak being ascended and the season of expedition.
-We notice that the success probabilities are pretty high for shorter mountain ranges and become lesser with increase in height.
+@fig-height-seasons shows the predicted success probability based on the height of the peak and the season of the expedition.
+Success probabilities are higher for shorter mountain ranges and decrease with increase in height.
 This might have to do with the fact that less higher peaks could be comparatively easier to summit.
-This, however, cannot be established as the trend as we see a non-pattern being followed for heights 6800 metres and above.
-We see that winter season expeditions have lower success probabilities, probably owning to the extreme cold conditions at higher altitudes.
-We also notice interesting trend as an interaction of these two factors.
-One such trend would be the success probabilities during different seasons for peaks between 5400 metres and 5749 metres are all comparatively closer to each other than compared to peaks between 7500 metres and 7849 metres in height.
-All seasons' success probabilities for peaks between 5400 metres and 5749 metres fall between 0.85 to 0.9 whereas for for peaks between 7500 metres and 7848 metres, the success probabilities go from around 0.35 in Winter to around 0.6 in Autumn or Spring.
+This, however, cannot be established as the trend as it is not consistent for peaks above 6800 meters.
+
+Winter expeditions have lower success probabilities, probably owning to the extreme cold conditions at higher altitudes.
+The interaction of these two factors provide an interesting insight.
+It is observed that the success probabilities during different seasons for peaks between 5400 metres and 5749 metres are all comparatively closer to each other than compared to peaks between 7500 metres and 7849 metres in height.
+All seasons' success probabilities for peaks between 5400 metres and 5749 metres fall between 0.85 to 0.9 whereas for for peaks between 7500 metres and 7848 metres, the success probabilities go from around 0.35 in winter to around 0.6 in autumn or spring.
 
 # Discussion {#sec-discussion}
-1) What is done in this paper? 
-2) What is something that we learn about the world?
+## Male dominance in mountaineering
+Mountaineering, as evident by the data, presents a significant gender gap, with men outnumbering women by nearly sixfold in expedition participation.
+@fig-counts shows us the counts of our independent variables and if we look at sex, we notice there are almost six times more male expedition members.
+This gender gap underscores the sport's male-dominated nature.
+
+The physical demands of mountaineering, requiring exceptional strength and endurance, align with biological differences between sexes, as men typically have greater muscle mass and strength.
+This may explain why men are more drawn to mountaineering due to their physical advantages.
+To have the same physical strength, a woman would need to work much harder.
+
+Despite the numerical difference, an interesting observation emerges from the analysis of success probabilities.
+@fig-model-coefficients and @fig-age-sex show how the success probabilities of men and women are not extremelyx different.
+While men have a slightly higher likelihood of reaching the summit, women show commendable success rates despite their lower representation in expeditions.
+
+This suggests that factors beyond sheer physical strength contribute to summit success, and one of those factors can be will power.
+As mentioned, women often face the reality of needing to exert greater effort due to inherent physical differences, coupled with the underrepresentation within mountaineering circles.
+The notable success rates among women despite these challenges highlight not only their physical strength but also underscore their exceptional mental resilience and inner motivation.
+
+These findings support the need for initiatives to promote gender diversity in mountaineering.
+Implementing programs and incentives tailored specifically for women can help overcome barriers, encourage more participation, and bring diverse perspectives and experiences to the mountaineering community.
+
+## Spring and Autumn are the best time for expeditions
+Autumn and spring, as transitional seasons, provide favorable conditions for mountaineering expeditions.
+The periods offer moderate temperatures and more stable weather patterns compared to the extremes of summer and winter.
+@fig-model-coefficients and @fig-height-seasons show the difference in probabilities of success by seasons and highlight highest chances of success in spring followed by autumn.
+
+Conversely, summer and winter present significant challenges for mountaineers.
+Summer brings the risks of intense heat and thunderstorms, increasing the likelihood of rockfall and other hazards while winter conditions are characterized by extreme cold, high winds, heavy snowfall, and increased avalanche danger, making ascent and descent more difficult and hazardous.
+As a result, winter is generally considered the least desirable season for mountaineering expeditions.
+
+In conclusion, autumn and spring are the optimal seasons for mountaineering expeditions, and season selection is a crucial factor in expedition planning.
+
+## Chances of success decline with increase in age
+Age is a key factor in mountaineering success, with success rates typically decreasing with increase in age.
+
+@fig-model-coefficients and @fig-age-sex show the impact of age on summit success.
+The impact of age on summit success is evident in the declining success rates with increase in age, influenced by factors such as reduced muscle mass, strength, and aerobic capacity associated with aging.
+Older climbers may also face challenges related to agility, flexibility, and recovery capacity, which can affect their performance on challenging terrain and recovery from strenuous exertion.
+
+Moreover, age-related health issues like cardiovascular problems, joint stiffness, and decreased bone density can further hinder older climbers' ability to meet the physical demands of mountaineering.
+These factors contribute to a decline in overall performance and increase the likelihood of encountering difficulties during expeditions.
+
+While age-related decline in mountaineering performance is noticeable, it is essential to recognize that success in mountaineering is not solely determined by physical capabilities.
+Experience, skill, preparation, and mindset also play crucial roles in summit success.
+Older climbers can compensate for physiological limitations through greater experience, strategic planning, and mental resilience, enabling them to pursue their mountaineering goals despite age-related challenges.
+
+In summary, age-related impacts on mountaineering success are significant, with success rates generally decreasing as climbers age.
+However, older climbers can overcome some of these challenges through experience, skill, strength training, and a strong mindset, allowing them to continue enjoying mountaineering well into their later years.
 
 ## Weaknesses and Limitations
+While my analysis offers valuable insights into the factors influencing a successful summit, it is important to acknowledge the weaknesses and limitations.
+
+Firstly, some success probabilities should be viewed with skepticism due to potential inaccuracies.
+Discrepancies in the counts of different categories may lead to imprecise success probabilities.
+For instance, the number of individuals attempting a solo ascent was significantly lower than those in group ascents.
+However, our model coefficients indicate for solo ascents was higher.
+This could create a misconception that solo ascents have better chances of success, potentially influencing more people to choose this option without considering other contributing factors.
+These probabilities may be influenced by the limited sample size of solo ascents, which lacks diverse data to account for confounding variables.
+There might be other factors which play a role like will power.
+Individuals opting for solo expeditions likely possess high levels of determination which can be a confounding factor playing into their success.
+
+Additionally, while we focused on focused on a few factors, there are numerous other variables that could impact the likelihood of a successful summit.
+Nationality and group size were two factors which were available in the raw data and I believed were factors which could have played a significant difference.
+These were later removed due to the lack of consistency in those variable’s.
+For instance, the way the number of members were reported differed from record to record.
+As the dataset also included the people summiting the other side of border peaks, some countries reported this number differently than others.
+For expeditions in Nepal, this was usually the number of foreigners listed on the expedition permit whereas in China, this was usually the number of non-hired members.
 
 ## Future Directions
+Moving forward, additional research is necessary to address the limitations discussed and improve the dataset to enhance our understanding of factors contributing to a successful summit.
+Including more diverse information can help uncover additional factors influencing a successful summit.
+Some of these factors that can be explored would be nationality of the person, attempt number, a metric for that could measure their level of preparedness and group size.
+
+Nationality can play a role as the cultural background and national resources can play a significant role in shaping an individual's climbing experience.
+Different countries may have different levels of access to training facilities, equipment, and support networks for climbers.
+Additionally, genetic differences between individuals may also play a role.
+
+Analysing the number of attempts made by climbers before a successful summit can reveal patterns related to perseverance, learning from past experiences, and adapting strategies.
+It can help determine whether repeated attempts increase success rates or reach a point of diminishing returns.
+
+Developing a metric to measure climbers' preparedness can quantify the various aspects of readiness, including physical fitness, technical skills, mental resilience, and logistical planning.
+Understanding the relationship between preparedness levels and summit success can guide training and preparation efforts.
+
+Group dynamics play a crucial role in mountaineering, affecting decision-making, safety, resource allocation, and overall team morale.
+Studying how group size influences summit success can provide insights into optimal team composition, leadership structures, and communication strategies.
+It can also highlight the trade-offs between the benefits of larger teams and the challenges of coordinating and managing them effectively.
 
@@
 
-1. Initial merging: The raw data from the expeditions and members datasets is merged based on a common identifier, expedition_id, consolidating information about expedition participants.
+1. Initial merging: The raw data from the `expeditions` and `members` datasets is merged based on a common identifier, `expedition_id`, consolidating information about expedition participants.
 
@@
 ## Analysis dataset
-Here is a glimpse of the dataset used for analysis
-
 ```{r}
@@
 ```
+@tbl-survey provides a glimpse of the dataset used for analysis.
 
 ## Model summary
-
-## Posterior distribution
 ```{r}
-#| label: fig-post_dist
-#| fig-cap: Posterior distribution for logistic regression model
 #| echo: false
@@
 #| message: false
+#| label: tbl-model-coefficients
+#| tbl-cap: Summary of the model
+
+# Extract coefficients
+coefficients <- broom::tidy(expeditions_model, conf.int = TRUE)
+coefficients |> kable(digits = 2)
+```
+@tbl-model-coefficients shows the coefficients for my Bayesian model along with the standard error and the 95% credible interval.
+The standard error (SE) is a measure of the precision with which a sample statistic estimates a population parameter.
+It quantifies the variability of sample statistics around the population parameter.
+
+## Posterior predictive check
+```{r}
+#| echo: false
+#| warning: false
+#| message: false
+#| label: fig-post
+#| fig-cap: Posterior distribution for logistic regression model
 
@@
   theme(legend.position = "bottom")
+
 ```
 
+```{r}
+#| echo: false
+#| warning: false
+#| message: false
+#| label: fig-prior-post
+#| fig-cap: Comparing prior distribution with posterior distribution
+
+library(broom)
+library(broom.mixed)
+library(knitr)
+library(modelsummary)
+library(purrr)
+library(rstanarm)
+library(testthat)
+library(tidyverse)
+
+posterior_vs_prior(expeditions_model) +
+  theme_minimal() +
+  theme(axis.text.x=element_text(hjust=1, vjust=0.5),
+        text=element_text(size=8)) +
+  guides(color = guide_legend(ncol = 6))+
+  theme(legend.position = "bottom") +
+  coord_flip()
+```
+@fig-post and @fig-prior-post are used to understand assess the quality of our model.
+
+@fig-post is the result of a posterior predictive check and is used to compare the actual outcome variable with simulations from the posterior distribution.
+@fig-prior-post is the result of comparing the posterior with the prior to see how much the estimates change once data are taken into account.
+Both these results support the model doing a good job of fitting the data.
+
+## Markov chain Monte Carlo (MCMC) Convergence Check
+```{r}
+#| echo: false
+#| eval: true
+#| message: false
+#| warning: false
+#| label: fig-trace
+#| fig-cap: "Trace plot"
+
+layout(matrix(1:6, ncol = 3))
+plot(expeditions_model, "trace")
+
+```
+
+```{r}
+#| echo: false
+#| eval: true
+#| message: false
+#| warning: false
+#| label: fig-rhat
+#| fig-cap: "Rhat plot"
+
+plot(expeditions_model, "rhat")
+```
+My model, modelled using `rstanarm` [@citeRstanarm] uses a sampling algorithm called Markov chain Monte Carlo (MCMC) to obtain samples from the posterior distributions of interest.
+@fig-trace and @fig-rhat are used to check for the existence of signs that the algorithm ran into issues.
+@fig-trace does not suggest anything out of the ordinary and everything in @fig-rhat is close to 1 suggesting no problem as well.
+
 # References

### 2024-04-19 paper/paper.qmd
Trace plots added

@@
 library(ggplot2)
-library(modelsummary)
 library(here)
@@
 library(arrow)
+library(rstanarm)
 
@@
 @fig-counts displays the counts for various variables I will be modelling.
-We see that there are significantly higher men than women.
+The plot shows there are significantly higher men than women.
 This shows the existence of a gender imbalance skewed towards men in mountaineering in the Himalayas.
 Additionally, most expedition members opt not to attempt solo ascents and fall within the middle-aged range.
-Lastly, we see the peak expedition seasons are the more pleasant autumn and spring over the extreme seasons like summer and winter.
+Lastly, it can be seen that the peak expedition seasons are the more pleasant autumn and spring over the extreme seasons like summer and winter.
 These count disparities should be considered when analysing success proportions.
@@
 @fig-proportions shows the proportion of expedition members who had a successful outcome or died during their expedition.
-For sex, we see that more men had successful expeditions compared to women.
+For sex, it is observed that more men had successful expeditions compared to women.
 Moreover, individuals who embark on solo expeditions have a higher mortality rate, potentially explaining the difference in the number of people attempting solo ascent observed earlier.
@@
 The use of moderately wide priors can also help regularize the model, preventing overfitting and providing more stable estimates, particularly when dealing with limited data.
+
+Model coefficients with their standard errors, posterior checks for the model and convergence checks for the model can be found in @sec-appendix
 
@@
 Mountaineering, as evident by the data, presents a significant gender gap, with men outnumbering women by nearly sixfold in expedition participation.
-@fig-counts shows us the counts of our independent variables and if we look at sex, we notice there are almost six times more male expedition members.
+@fig-counts shows the counts of the independent variables and it is observed that there are almost six times more male expedition members.
 This gender gap underscores the sport's male-dominated nature.
@@
 For instance, the number of individuals attempting a solo ascent was significantly lower than those in group ascents.
-However, our model coefficients indicate for solo ascents was higher.
+However, my model coefficients indicate for solo ascents was higher.
 This could create a misconception that solo ascents have better chances of success, potentially influencing more people to choose this option without considering other contributing factors.
@@
 
-Additionally, while we focused on focused on a few factors, there are numerous other variables that could impact the likelihood of a successful summit.
+Additionally, while I focused on focused on a few factors, there are numerous other variables that could impact the likelihood of a successful summit.
 Nationality and group size were two factors which were available in the raw data and I believed were factors which could have played a significant difference.
@@
 ## Future Directions
-Moving forward, additional research is necessary to address the limitations discussed and improve the dataset to enhance our understanding of factors contributing to a successful summit.
+Moving forward, additional research is necessary to address the limitations discussed and improve the dataset to enhance the understanding of factors contributing to a successful summit.
 Including more diverse information can help uncover additional factors influencing a successful summit.
@@
 
-# Appendix
+# Appendix {#sec-appendix}
 
@@
 
-library(bayesplot)
-
 pp_check(expeditions_model) +
   theme(legend.position = "bottom")
-
 ```
@@
 
-library(broom)
-library(broom.mixed)
-library(knitr)
-library(modelsummary)
-library(purrr)
-library(rstanarm)
-library(testthat)
-library(tidyverse)
-
 posterior_vs_prior(expeditions_model) +
   theme_minimal() +
-  theme(axis.text.x=element_text(hjust=1, vjust=0.5),
-        text=element_text(size=8)) +
-  guides(color = guide_legend(ncol = 6))+
+  theme(
+    axis.text.x = element_text(hjust = 1, vjust = 0.5),
+    text = element_text(size = 8)
+  ) +
+  guides(color = guide_legend(ncol = 6)) +
   theme(legend.position = "bottom") +
@@
 ```
-@fig-post and @fig-prior-post are used to understand assess the quality of our model.
+@fig-post and @fig-prior-post are used to understand assess the quality of my model.
 
@@
 #| warning: false
-#| label: fig-trace
+#| label: fig-trace1
 #| fig-cap: "Trace plot"
+#| fig-subcap: ["Intercept", "sexM", "soloTRUE", "seasonsSpring", "seasonsSummer", "seasonsWinter", "age_rangeUnder 18",  "age_range31-40", "age_range41-50", "age_range51-60", "age_range61-70", "age_range71-80", "age_range81-90"]
+#| layout-ncol: 3
 
-layout(matrix(1:6, ncol = 3))
-plot(expeditions_model, "trace")
+plot(expeditions_model, "trace", "(Intercept)")
+plot(expeditions_model, "trace", "sexM")
+plot(expeditions_model, "trace", "soloTRUE")
+plot(expeditions_model, "trace", "seasonsSpring")
+plot(expeditions_model, "trace", "seasonsSummer")
+plot(expeditions_model, "trace", "seasonsWinter")
+plot(expeditions_model, "trace", "age_rangeUnder 18")
+plot(expeditions_model, "trace", "age_range31-40")
+plot(expeditions_model, "trace", "age_range41-50")
+plot(expeditions_model, "trace", "age_range51-60")
+plot(expeditions_model, "trace", "age_range61-70")
+plot(expeditions_model, "trace", "age_range71-80")
+plot(expeditions_model, "trace", "age_range81-90")
+```
+```{r}
+#| echo: false
+#| eval: true
+#| message: false
+#| warning: false
+#| label: fig-trace2
+#| fig-cap: "Trace plot"
+#| fig-subcap: ["height_range5750 - 6099", "height_range6100 - 6449", "height_range6450 - 6799", "height_range6800 - 7149", "height_range7150 - 7499", "height_range7500 - 7849", "height_range7850 - 8199", "height_range8200 - 8549", "height_range8550 - 8900"]
+#| layout-ncol: 3
 
+plot(expeditions_model, "trace", "height_range5750 - 6099")
+plot(expeditions_model, "trace", "height_range6100 - 6449")
+plot(expeditions_model, "trace", "height_range6450 - 6799")
+plot(expeditions_model, "trace", "height_range6800 - 7149")
+plot(expeditions_model, "trace", "height_range7150 - 7499")
+plot(expeditions_model, "trace", "height_range7500 - 7849")
+plot(expeditions_model, "trace", "height_range7850 - 8199")
+plot(expeditions_model, "trace", "height_range8200 - 8549")
+plot(expeditions_model, "trace", "height_range8550 - 8900")
 ```
@@
 My model, modelled using `rstanarm` [@citeRstanarm] uses a sampling algorithm called Markov chain Monte Carlo (MCMC) to obtain samples from the posterior distributions of interest.
-@fig-trace and @fig-rhat are used to check for the existence of signs that the algorithm ran into issues.
-@fig-trace does not suggest anything out of the ordinary and everything in @fig-rhat is close to 1 suggesting no problem as well.
+@fig-trace1, @fig-trace2 and @fig-rhat are used to check for the existence of signs that the algorithm ran into issues.
+@fig-trace1 and @fig-trace2 does not suggest anything out of the ordinary and everything in @fig-rhat is close to 1 suggesting no problem as well.
 

### 2024-04-19 scripts/00-install_packages.R

@@ -14,7 +14,6 @@ packages <- c(
   "dplyr",
   "knitr",
   "ggplot2",
-  "modelsummary",
   "here",
   "kableExtra",
   "arrow"

### 2024-04-19 paper/paper.qmd
Final changes

@@
 abstract: "This study analyses the relationship between various demographic, environmental and geographic factors and the probability of success in attempting to summit a peak.
-Data from Himalayan expeditions in Nepal from 1905 through Spring 2019 is used and a Bayesian Logistic Regression model is leveraged to analyse the trends and factors influencing a successful summit.
+Data from Himalayan expeditions in Nepal from 1905 through Spring 2019 is used and a Bayesian Logistic Regression model is utilised to analyse the trends and factors influencing a successful summit.
 Young age, being male and embarking on the expedition in spring or autumn are some factors that increase one's chances of a successful summit.
@@
 Amidst the allure of conquering majestic peaks, a critical question arises: what factors contribute to the likelihood of a successful summit attempt?
-This paper delves into precisely this.
+This paper tries to unravel precisely this.
 
@@
 
+  # Custom labels for "sex" variable
+  sex_labels <- c("Female", "Male")
+
   # Create a bar plot
-  ggplot(data, aes_string(x = x)) +
+  p <- ggplot(data, aes_string(x = x)) +
     geom_bar(fill = "lightblue") +
@@
     theme(axis.text.x = element_text(angle = 45, hjust = 1))
+
+  # If x is "sex", set custom labels
+  if (x == "sex") {
+    p <- p + scale_x_discrete(labels = sex_labels)
+  }
+
+  return(p)
 }
@@
 
+  # Custom labels for "sex" variable
+  sex_labels <- c("Female", "Male")
+
   p <- ggplot(plot_data, aes(x = x_var, y = proportion, fill = factor(success))) +
@@
     theme(axis.text.x = element_text(angle = 45, hjust = 1))
+
+  # If x is "sex", set custom labels
+  if (x_var == "sex") {
+    p <- p + scale_x_discrete(labels = sex_labels)
+  }
 
@@
   facet_grid(sex ~ .) +
+  scale_fill_discrete(labels = c("Female", "Male")) +
   theme_minimal()
@@
 The sharp decline in the 81-90 age category may be attributed to the smaller number of expedition members in that age group, resulting in less reliable outcomes.
+
+\newpage
 
@@
 
+\newpage
+
 # Discussion {#sec-discussion}
@@
 
-The physical demands of mountaineering, requiring exceptional strength and endurance, align with biological differences between sexes, as men typically have greater muscle mass and strength.
+The physical demands of mountaineering, requiring exceptional strength and endurance [@Strength], align with biological differences between sexes, as men typically have greater muscle mass and strength [@Sex].
 This may explain why men are more drawn to mountaineering due to their physical advantages.
@@
 Despite the numerical difference, an interesting observation emerges from the analysis of success probabilities.
-@fig-model-coefficients and @fig-age-sex show how the success probabilities of men and women are not extremelyx different.
+@fig-model-coefficients and @fig-age-sex show how the success probabilities of men and women are not extremely different.
 While men have a slightly higher likelihood of reaching the summit, women show commendable success rates despite their lower representation in expeditions.
@@
 The periods offer moderate temperatures and more stable weather patterns compared to the extremes of summer and winter.
-@fig-model-coefficients and @fig-height-seasons show the difference in probabilities of success by seasons and highlight highest chances of success in spring followed by autumn.
+@fig-model-coefficients and @fig-height-seasons show the difference in probabilities of success by seasons and highlight higher chances of success in spring and autumn.
+@Season provides tips for summit success when climbing Mt.
+Everest and suggest that the best time to climb it is in the Autumn/ Fall season.
+This is in-line with our findings.
 
@@
 
-In conclusion, autumn and spring are the optimal seasons for mountaineering expeditions, and season selection is a crucial factor in expedition planning.
+In conclusion, autumn and spring are the optimal seasons for mountaineering expeditions, and season selection is an important factor in expedition planning.
 
@@
 
+@Age highlight some of the same issues and factors when analysing endurance performance in masters athletes.
+
 While age-related decline in mountaineering performance is noticeable, it is essential to recognize that success in mountaineering is not solely determined by physical capabilities.
-Experience, skill, preparation, and mindset also play crucial roles in summit success.
+Experience, skill, preparation, and mindset also play an important roles in summit success.
 Older climbers can compensate for physiological limitations through greater experience, strategic planning, and mental resilience, enabling them to pursue their mountaineering goals despite age-related challenges.
@@
 However, older climbers can overcome some of these challenges through experience, skill, strength training, and a strong mindset, allowing them to continue enjoying mountaineering well into their later years.
+
+\newpage
 
@@
 
-Group dynamics play a crucial role in mountaineering, affecting decision-making, safety, resource allocation, and overall team morale.
+Group dynamics play an important role in mountaineering, affecting decision-making, safety, resource allocation, and overall team morale.
 Studying how group size influences summit success can provide insights into optimal team composition, leadership structures, and communication strategies.
@@
 
+# Posterior check
 pp_check(expeditions_model) +
@@
 #| echo: false
-#| eval: true
 #| message: false
@@
 
+# Split plots as all at once is unreadable
 plot(expeditions_model, "trace", "(Intercept)")
@@
 
+\newpage
+
 # References

### 2024-04-19 scripts/05-predictions.R
Update 05-predictions.R preamble

@@ -1,5 +1,5 @@
 #### Preamble ####
-# Purpose: Creates prediction dataset to use in Results of the paper
+# Purpose: Creates prediction dataset to use in the Results section of the paper
 # Author: Kaavya Kalani
 # Date: 18 April 2024
 # Contact: kaavya.kalani@mail.utoronto.ca

### 2024-04-29 paper/paper.qmd

@@
 The Himalayan Database is a compilation of records for all expeditions that have climbed in the Nepal Himalaya.
-The database is based on the expedition archives of Elizabeth Hawley, a longtime journalist based in Kathmandu, and it is supplemented by information gathered from books, alpine journals and correspondence with Himalayan climbers.
+The database is based on the expedition archives of Elizabeth Hawley, a long-time journalist based in Kathmandu, and it is supplemented by information gathered from books, alpine journals and correspondence with Himalayan climbers.
 
@@
 The three datasets I considered included information about all peaks in Nepal, all expeditions on those peaks and all members on those expeditions.
-I then combines the data from these three datasets to form the main analysis dataset.
+I then combine the data from these three datasets to form the main analysis dataset.
 
@@
 
-\newpage
-
 # Model {#sec-model}
@@
 
-\newpage
-
 ```{r}
@@
 It is observed that the success probabilities during different seasons for peaks between 5400 metres and 5749 metres are all comparatively closer to each other than compared to peaks between 7500 metres and 7849 metres in height.
-All seasons' success probabilities for peaks between 5400 metres and 5749 metres fall between 0.85 to 0.9 whereas for for peaks between 7500 metres and 7848 metres, the success probabilities go from around 0.35 in winter to around 0.6 in autumn or spring.
-
-\newpage
+All seasons' success probabilities for peaks between 5400 metres and 5749 metres fall between 0.85 to 0.9 whereas for peaks between 7500 metres and 7848 metres, the success probabilities go from around 0.35 in winter to around 0.6 in autumn or spring.
 
@@
 While age-related decline in mountaineering performance is noticeable, it is essential to recognize that success in mountaineering is not solely determined by physical capabilities.
-Experience, skill, preparation, and mindset also play an important roles in summit success.
+Experience, skill, preparation, and mindset also play important roles in summit success.
 Older climbers can compensate for physiological limitations through greater experience, strategic planning, and mental resilience, enabling them to pursue their mountaineering goals despite age-related challenges.
@@
 However, older climbers can overcome some of these challenges through experience, skill, strength training, and a strong mindset, allowing them to continue enjoying mountaineering well into their later years.
-
-\newpage
 
@@
 Nationality and group size were two factors which were available in the raw data and I believed were factors which could have played a significant difference.
-These were later removed due to the lack of consistency in those variable’s.
+These were later removed due to the lack of consistency in those variables.
 For instance, the way the number of members were reported differed from record to record.
@@
 It can also highlight the trade-offs between the benefits of larger teams and the challenges of coordinating and managing them effectively.
+
+\newpage
+
+\appendix
 
